### 2020-10-13 reports/rmarkdown/01_data_pre_processing.Rmd
adds data pre-processing and QC

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/data/" #Path to the data

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# Load datasets
dataset_names <- c("ASV_tissue_B2.csv",
                     "ASV_tissue_B1.csv",
                     "ASV_CVL_V2_B1.csv",
                     "ASV_CVL_V2_B2.csv",
                     "ASV_CVL_V2_C.csv",
                     "ASV_CVL_V3.csv")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  sps <- apply( temp[,6:7] ,1,function(x) paste(x,collapse ="_"))
  not_complete <- grepl("_NA",sps)

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")

  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)

#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)

# datasets <- datasets[3] #in case you wanna test in 1 dataset

# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
```

# Merging, renaming and correcting batch effects on datasets

```{r, echo = FALSE}
#No batch correction needed in tissue_V3 dataset, only add it to the other samples
datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_B2.csv"]],datasets[["ASV_tissue_B1.csv"]] )

datasets <- datasets[-c(1,2)]
lapply(datasets,dim)

#Batch correction for the sample CVL_V2
batch1 <- cbind( datasets[["ASV_CVL_V2_B1.csv"]],datasets[["ASV_CVL_V2_B2.csv"]] )
batch1 <- t( log2( t(batch1) / colSums(batch1) * 1000 + 1 ))
batch2 <- datasets[["ASV_CVL_V2_C.csv"]]
batch2 <- t( log2( t(batch2) / colSums(batch2) * 1000 + 1 ))

common_samples <- sort(colnames(batch1) [ colnames(batch1) %in% colnames(batch2) ])
colnames(batch1) <- paste0("boston_",colnames(batch1))
colnames(batch2) <- paste0("CTMR_",colnames(batch2))

merged <- cbind(batch1,batch2)
batch <- factor(sub("_.*","",colnames(merged)))
common_samples2 <- sort(colnames(merged) [ sub(".*_","",colnames(merged)) %in% common_samples])

nonzeros <- rowSums(merged)>0
merged <- merged[nonzeros,]

PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )
```

# Batch correction with different approaches

```{r, echo = FALSE}
mypar(3,3)
PC_raw <- prcomp(round(t(batch1[nonzeros, paste0("boston_",common_samples)]),4) )

#COMBAT
mod0 <- model.matrix(~1,data = data.frame(colnames(merged),row.names = colnames(merged)))
corrected_combat <- ComBat(dat=merged, batch=batch, mod=mod0)

#mnnCorrect
corrected_mnnCorrect <- batchelor::mnnCorrect( batch2[nonzeros,],
                                               batch1[nonzeros,] ,
                                               cos.norm.out = F,cos.norm.in = F,
                                         restrict = list((1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ],
                                                         (1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ]))
corrected_mnnCorrect <- corrected_mnnCorrect@assays@data$corrected

#regressBatches
corrected_regressBatches <- batchelor::regressBatches( batch1[nonzeros,] ,
                                         batch2[nonzeros,],
                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
corrected_regressBatches <- corrected_regressBatches@assays@data$corrected

#rescaleBatches
corrected_rescaleBatches <- batchelor::rescaleBatches( batch1[nonzeros,] ,
                                         batch2[nonzeros,],
                                         restrict = list((1:ncol(batch1)) [ colnames(batch1) %in% common_samples2 ],
                                                         (1:ncol(batch2)) [ colnames(batch2) %in% common_samples2 ]) )
corrected_rescaleBatches <- corrected_rescaleBatches@assays@data$corrected
```

# Batch correction with Autoencoder

```{r, echo = FALSE}
library(keras)
library(caret)

xtrain <- t(batch1[nonzeros,paste0("boston_",sort(common_samples))])
xtrain <- xtrain/max(xtrain)
ytrain <- t(batch2[nonzeros,paste0("CTMR_",sort(common_samples))])
ytrain <- ytrain/max(ytrain)
dim(xtrain)

input_size = dim(xtrain)[2]
latent_size = 10
print(input_size)

x_train = array_reshape(xtrain, dim=c(dim(xtrain)[1], input_size))
y_train = array_reshape(ytrain, dim=c(dim(ytrain)[1], input_size))

print(dim(x_train))
print(dim(y_train))

if(file.exists(paste0(PATH,"../results/encoder.h5")) & 
  file.exists(paste0(PATH,"../results/decoder.h5")) & 
  file.exists(paste0(PATH,"../results/aen.h5")) ){
  encoder <- load_model_hdf5(paste0(PATH,"../results/encoder.h5"))
  decoder <- load_model_hdf5(paste0(PATH,"../results/decoder.h5"))
  aen <- load_model_hdf5(paste0(PATH,"../results/aen.h5"))

} else {
  enc_input = layer_input(shape = input_size)
  enc_output = enc_input %>% 
    layer_dense(units=200, activation = "relu") %>% 
    layer_activation_leaky_relu() %>% 
    layer_dense(units=latent_size,activation = "sigmoid") %>% 
    layer_activation_leaky_relu()

  encoder = keras_model(enc_input, enc_output)
  summary(encoder)

  dec_input = layer_input(shape = latent_size)
  dec_output = dec_input %>% 
    layer_dense(units=200, activation = "relu") %>% 
    layer_activation_leaky_relu() %>% 
    layer_dense(units = input_size, activation = "sigmoid") %>% 
    layer_activation_leaky_relu()

  decoder = keras_model(dec_input, dec_output)
  summary(decoder)

  aen_input = layer_input(shape = input_size)
  aen_output = aen_input %>% 
    encoder() %>% 
    decoder()

  aen = keras_model(aen_input, aen_output)
  summary(aen)

  aen %>% compile(optimizer="rmsprop", loss="binary_crossentropy")

  aen %>% fit( y_train , x_train , epochs=2000, batch_size=round(dim(y_train)[1]/2) )

  save_model_hdf5(encoder, paste0(PATH,"../results/encoder.h5") )
  save_model_hdf5(decoder, paste0(PATH,"../results/decoder.h5"))
  save_model_hdf5(aen, paste0(PATH,"../results/aen.h5"))
}

encoded_imgs = encoder %>% predict(y_train)
decoded_imgs = decoder %>% predict(encoded_imgs)

# plot( x_train , decoded_imgs , cex=.3)
# abline(0,1,col="red")
# plot( x_train, y_train , cex=.3 )
# abline(0,1,col="red")
```

# Plotting the batch-corrected data for comparisson

```{r, echo = FALSE,fig.height=7*2,fig.width=7}
layout(matrix(c(1,1,1,1,     
                2,2,2,3,
                4,4,4,5,
                6,6,6,7,
                8,8,8,9,
                10,10,10,11,
                12,12,12,13
                ) , ncol= 4,byrow = T),heights = c(1))

fig_letters <- letters
par(mar=c(3,3,3,1))
pc_var <- PC_raw$sdev^2 / sum(PC_raw$sdev^2) * 100
barplot(pc_var,ylim=c(0,60),las=2,names.arg = paste0("PC",1:length(pc_var)),
        border = NA,col="grey20",ylab="% variance explained")
abline(h=0,lwd=2)
add_letter(fig_letters); fig_letters <- fig_letters[-1]
legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="uncorrected") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(batch2[nonzeros, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="ComBat") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_combat[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="mnnCorrect") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_mnnCorrect[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="regressBatches") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_regressBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="rescaleBatches") ; axis(2)
abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(corrected_rescaleBatches[, paste0("CTMR_",common_samples)]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

plot(c(1,20),range(PC_raw$x),type="n",ylim=c(-.2,1.2),yaxs="i",frame=F,axes=F,ylab="PC embbeddings",xlab="",main="AutoEncoder") ; axis(2)
# abline(h=0,col="grey90",lty=1)
lines(c(par("usr")[1],par("usr")[1]),c(par("usr")[3],par("usr")[4]),xpd=T)
b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:20){
  abline(v=i,col="grey90",lty=1)
  x <- range(PC_raw$x[,i])
  points(rep(i-.2,nrow(PC_raw$x)), (PC_raw$x[,i]-x[1]) / (x[2]-x[1]),cex=.5, pch=1,col= "firebrick" )
  for(j in 1:nrow(b2_projection)){ lines( c(i-.2,i+.2) , c((PC_raw$x[j,i]-x[1]) / (x[2]-x[1]), (b2_projection[j,i]-x[1])/(x[2]-x[1])) , col="grey50") }
  points(rep(i+.2,nrow(b2_projection)), (b2_projection[,i]-x[1])/(x[2]-x[1]),cex=.5, pch=1,col= "navy" )
  text(i,par("usr")[3],labels = paste0("PC",i),xpd=T,srt=90,adj=c(1,.5))
  # wt <- t.test(PC_raw$x[,i],b2_projection[,i], paired = T)
  # text(i,par("usr")[4],labels = wt$p.value,xpd=T,srt=90,adj=c(0,.5))
}
add_letter(fig_letters); fig_letters <- fig_letters[-1]
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", xlim=c(-20,15),ylim=c(-15,15))
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey50") }
points(b2_projection[,1:2], cex=.5, pch=2, col= "navy")

```

As we can observe above, when using enough epochs (between 1000 and 2000), the autoencoder fits the data and learns the structure of the data.

```{r, echo = FALSE}
batch2_all <- t(batch2[nonzeros,])
batch2_all <- batch2_all/max(batch2_all)

batch2_encoded_data <- encoder %>% predict(batch2_all)
batch2_decoded_data <- decoder %>% predict(batch2_encoded_data)
batch2_decoded_data <- t(round(t(t(batch2_decoded_data*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4))
rownames(batch2_decoded_data) <- colnames(batch2_all) ; colnames(batch2_decoded_data) <- rownames(batch2_all)

mypar()
plot(PC_raw$x[,1:2], cex=.5, pch=1, col= "firebrick", main="AutoEncoder",xlim=c(-20,15),ylim=c(-15,15),type="n")
b2_batch_projection <- scale(t(batch2_decoded_data), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
b1_projection <- scale(round(t(batch1[nonzeros,]),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation

b2_projection <- scale(round(t(t(decoded_imgs*max(t(batch1[nonzeros,paste0("boston_",sort(common_samples))])))),4), PC_raw$center, PC_raw$scale) %*% PC_raw$rotation
for(i in 1:nrow(b2_projection)){ lines( c(PC_raw$x[i,1],b2_projection[i,1]) , c(PC_raw$x[i,2], b2_projection[i,2]) , col="grey20") }

points(b2_batch_projection[,1:2], cex=.8, pch=15, col= "navy")
points(b1_projection[,1:2], cex=.8, pch=16, col= "firebrick")
legend("topright",legend = c("batch1","batch2"),col = c("firebrick","navy"),pch=c(15,16),bty = "n")

samples_not_batch1 <- sort(colnames(batch2_decoded_data) [! colnames(batch2_decoded_data) %in% common_samples2])
datasets[["ASV_CVL_V2"]] <- cbind( batch1[nonzeros,] , batch2_decoded_data[,samples_not_batch1] )
datasets <- datasets[-c(1,2,3)]
colnames(datasets[["ASV_CVL_V2"]]) <- sub(".*_","",colnames(datasets[["ASV_CVL_V2"]]))
lapply(datasets,dim)
lapply(datasets,colnames)

common_microbes <- unique(unlist(lapply(datasets,rownames)))
all_samples <- rownames(metadata)

datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)

datasets <- lapply(datasets,function(x){
  x <- x[,sort(colnames(x))]
  return(x)
})

datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)

write.csv( datasets[["ASV_CVL_V2"]] , paste0(PATH,"../results/ASV_CVL_V2_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_tissue_V3"]] , paste0(PATH,"../results/ASV_tissue_V3_normalized_batch_corrected.csv"),row.names = T )
write.csv( datasets[["ASV_CVL_V3.csv"]] , paste0(PATH,"../results/ASV_CVL_V3_normalized_batch_corrected.csv"),row.names = T )

```

```{r, echo = FALSE}
trx <- read.csv(paste0(PATH,"/TRX_rawcounts.csv"),row.names = 1)
trx <- rowsum(trx[,-c(1,2)],trx$symbol)
colnames(trx) <- sub("ML","",colnames(trx))
trx  <- trx[ rownames(trx) != "" , ]
colnames(trx) <- as.character(metadata$ID)[ match(colnames(trx) , as.character(metadata$PatID) ) ]
trx <- trx[ rowSums(trx > 5) >= 3 , ]
trx <- t( log2( t( trx ) / colSums(trx) * 1e6 + 1 ) )
write.csv( trx , paste0(PATH,"../results/Tissue_RNAseq_V3_normalized.csv"),row.names = T )
```