### 2020-10-13 reports/rmarkdown/02_data_quality_control.Rmd

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(sva)
library(batchelor)
source("~/repos/niceRplots/R/plotting_functions.R")
source("~/repos/niceRplots/R/add_fig_label.R")
source("~/repos/niceRplots/R/helper_functions.R")

```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data

#color palettes
pal <- RColorBrewer::brewer.pal(8,"Set1") #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# Load datasets
dataset_names <- c("Tissue_RNAseq_V3_normalized",
                     "ASV_tissue_V3_normalized_batch_corrected",
                     "ASV_CVL_V3_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_batch_corrected")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"results/",x,".csv"),row.names = 1)
  return(as.matrix(x)) })
names(datasets) <- dataset_names

#fill all samples with 0s to make plotting easier
datasets_all_samples <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = nrow(x),
                 ncol = nrow(metadata),
                 dimnames = list(rownames(x),rownames(metadata)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets_all_samples,dim)

```

# Calculate QC metrics

```{r, echo = FALSE}
counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
counts[is.na(counts)] <- 0
detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
detected_counts[is.na(detected_counts)] <- 0
rownames(detected_counts) <- paste0(rownames(detected_counts),">0")

mypar(mar=c(2,20,2,1))
barlist( data = rbind( counts , detected_counts),
         genes = c(rownames(counts),rownames(detected_counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)

sum((colSums(counts == 0)>=3)*1)
```

# Calculate QC metrics

```{r, echo = FALSE}
counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
counts[is.na(counts)] <- 0
detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
detected_counts[is.na(detected_counts)] <- 0
rownames(detected_counts) <- paste0(rownames(detected_counts),">0")

mypar(mar=c(2,20,2,1))
barlist( data = rbind( counts , detected_counts),
         genes = c(rownames(counts),rownames(detected_counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)>=3)*1,"red","grey" ),draw_mean_lines=F)

sum((colSums(counts == 0)>=3)*1)
```