### 2020-10-28 reports/rmarkdown/00_knit_all_reports.Rmd
adds the report for comparing microbiome datasets

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r, echo = T}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325" #Path to the data
```

```{r, echo = T}
rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/01_data_pre_processing.Rmd"),
                  output_format = "pdf_document")
```

```{r, echo = T}
rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/02_data_quality_control.Rmd"),
                  output_format = "pdf_document")
```

```{r, echo = T}
dataset_names <- c("ASV_tissue_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V3_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_batch_corrected.csv",
                   "ASV_CVL_V2_normalized_NOT_batch_corrected.csv")

for(i in dataset_names){
  rmarkdown::render(input = paste0(PATH,"/reports/rmarkdown/03_individual_reports.Rmd"),
                    output_format = "pdf_document",
                    params = list(dataset=i),
                    output_file = paste0(PATH,"/reports/rmarkdown/report_",sub(".csv","",i)))
}
```