### 2020-11-27 reports/rmarkdown/03_integrated_datasets.Rmd
updates integrated report

---
title: "`r params$project_title`"
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
params:
  organisation: Karolinska Institutet
  pi_mail: prin.cipal@ki.se
  pi_name: Prin Cipal
  project_title: Broliden_5325
  redmine_issue: '#5325'
  requester_mail: kristina.broliden@ki.se
  requester_name: Kristina Broliden
  staff_mail: supp.ortstaff@nbis.se
  staff_name: Supp Ortstaff
  staff_web: https://nbis.se/about/staff/supp-ortstaff/
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 10,
                      results    = "hide",
                      message    = FALSE,
                      warning    = FALSE)
```

#Load libraries and other scripts

```{r, echo = FALSE}
library(igraph)
library(rafalib)
library(fgsea)
library(scales)
remotes::install_github("czarnewski/niceRplots",upgrade = "always",force = T)
library(niceRplots)
```

#Defining some variables for the analysis

```{r, echo = FALSE}
PATH <- "~/Desktop/NBIS/SMS_Projects/broliden_5325/" #Path to the data

#color palettes
pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)

#Graph construction  
fct <- .25 # FC threshold for differential expression
pvt <- 0.01 # Pvalue threshold for differential expression
min_pct <- .05 # minimun level of detected bacteria in each sample group

```

# Loading data and metadata

```{r, echo = FALSE}
# Load metadata
metadata <- read.csv(paste0(PATH,"/data/Clinical_visit_2_3_updatesept28.csv"))
rownames(metadata) <- as.character(metadata$ID)

# Load datasets
dataset_names <- c("Tissue_RNAseq_V3_normalized",
                     "ASV_tissue_V3_normalized_batch_corrected",
                     "ASV_CVL_V3_normalized_batch_corrected",
                     "ASV_CVL_V2_normalized_NOT_batch_corrected")

#import datasets and return matrix with taxa-level resolution
datasets <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"results/",x,".csv"),row.names = 1)
  return(as.matrix(x)) })
names(datasets) <- dataset_names

#fill all samples with 0s to make plotting easier
datasets_all_samples <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = nrow(x),
                 ncol = nrow(metadata),
                 dimnames = list(rownames(x),rownames(metadata)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets_all_samples,dim)
```

# Calculate QC metrics

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=5}
counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
counts[is.na(counts)] <- 0
detected_counts <- t(as.data.frame( lapply(datasets_all_samples,function(x) colSums( x > 0 )) ))
detected_counts[is.na(detected_counts)] <- 0
rownames(detected_counts) <- paste0(rownames(detected_counts),">0")

mypar(2,1,mar=c(2,10,2,1))
barlist( data = rbind( counts ),main = "total counts",
         genes = c(rownames(counts)), labels = sub("_norm.*","",rownames(counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
add_letter("a")

barlist( data = rbind( detected_counts),main = "detected features",
         genes = c(rownames(detected_counts)),labels = sub("_norm.*","",rownames(counts)),
         clustering = rep(1,nrow(metadata)),
         col=ifelse( (colSums(counts == 0)==3)*1,"red", ifelse( (colSums(counts == 0)==2)*1,"orange3","grey" ) ),draw_mean_lines=F)
add_letter("b")

sum((colSums(counts == 0)>=3)*1)
colnames(counts)[(colSums(counts == 0)==3)]
paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")

metadata[colnames(counts)[(colSums(counts == 0)>=3)],]
```

**Figure 1.** Comparative barplot for the **a)** total counts and **b)** number of non-zero detected features (genes / bacteria) for each of the sequencing datasets. Samples are ordered alfabetically according to the patient ID. Samples that are present in exactly two datasets are shown in orange (`r paste0(colnames(counts)[(colSums(counts == 0)==2)],collapse = ",")`). Samples that are present in exactly two datasets are shown in red (`r paste0(colnames(counts)[(colSums(counts == 0)==3)],collapse = ",")`).

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- rbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0)>=2 ,  ]

zscore <- t(apply(all_microbiome,1,function(i){scale(i,T,T)}))
zscore[ zscore > 6 ] <- 6
zscore[ zscore < -6 ] <- -6
rownames(zscore) <- rownames(all_microbiome); colnames(zscore) <- colnames(all_microbiome)

#Compute sample-wise correlations
cors <- cor(zscore, method = "pearson")

#Graph construction  
k <- 5 # k for KNN
P <- 0.4# prunning threshold of SNN

#Compute KNN graph from correlation matrix
knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Compute SNN graph from KNN
jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))

#Prune SNN graph based on the Jaccard index
pSNN <- jSNN
pSNN <- pSNN / max(pSNN)

# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="SNN jaccard",xlim=c(0,1))
# abline(v=P,col="red",lty=2,xpd=F)
# 
# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="correlation coeficient",xlim=c(-1,1))
# abline(v=0,col="red",lty=1,xpd=F)

pSNN[ pSNN < P ] <- 0
pSNN[ pSNN > P ] <- cors[pSNN > P]
pSNN[ pSNN < 0 ] <- 0

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)

mypar(2,2,mar=c(0,0,2,10))

set.seed(1)
# l <- layout_with_graphopt(g,niter = 1000)
l <- layout_with_fr(g,niter=3000,start.temp=30)

cl <- igraph::cluster_louvain(g)
# cl <- igraph::walktrap.community(g)
metadata$joint_clustering <- cl$membership

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)

```

**Figure 1b.** Sample embedding of `r k`-SNN graph clustered using Louvain based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=4}
myletters <- letters

mypar(1,2,mar=c(2,15,1,1))

for (m in c("ASV_tissue_V3_normalized_batch_corrected",
            "ASV_CVL_V3_normalized_batch_corrected",
            "ASV_CVL_V2_normalized_NOT_batch_corrected")){

  all_microbiome <- datasets_all_samples[[m]]

  datasets <- factor(metadata$joint_clustering)
  # datasets <- datasets[ colSums(all_microbiome)>0 ]
  # all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

  NN <- min(table(datasets))

  res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
  for(i in levels(datasets)){
    for(j in rownames(all_microbiome) ){

      set.seed(1)
      a <- c(sample(all_microbiome[j,datasets == i],NN))
      set.seed(1)
      b <- sample(all_microbiome[j,datasets != i],NN)

      perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
      perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

      temp <- wilcox.test(x=a, y=b)

      fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

      res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                            c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
      colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
    }
  }
  res <- res[-1,]
  res$pvalue <- as.numeric(res$pvalue)
  res$perc.1 <- as.numeric(res$perc.1)
  res$perc.2 <- as.numeric(res$perc.2)
  res$perc.diff <- res$perc.1 - res$perc.2

  res$fc <- as.numeric(res$fc)
  res$FDR <- p.adjust(res$pvalue)
  # res <- res[order(res$pvalue),]
  res <- res[abs(res$fc) > fct,]
  res <- res[abs(res$pvalue) < 0.05,]
  res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
  dim(res)

  ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

  plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
            cex.row = .8,cex.col = .8,srt = 0,main = m)
  add_letter(myletters); myletters <- myletters[-1]

  barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
            cex.axis = .8,srt = 0,main = m)
  add_letter(myletters); myletters <- myletters[-1]

}
```

**Figure 2.** Differential bacterial abundance across joint clustered samples and compared across microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,2,mar=c(2,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)
add_letter("a")
barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.axis = .8,srt = 0)
add_letter("b")
```

**Figure 2.** Differential bacterial abundance across microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- factor(c( rep("CVL_v3",121) , rep("CVL_v2",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,2,mar=c(2,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)
add_letter("a")
barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.axis = .8,srt = 0)
add_letter("b")
```

**Figure 2.** Differential bacterial abundance across CVL2 and CVL3 microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]])
datasets <- factor(c( rep("tissue_v3",121) , rep("CVL_v3",121) ))
datasets <- datasets[ colSums(all_microbiome)>0 ]
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))

res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,2,mar=c(2,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .8,cex.col = .8,srt = 0)
add_letter("a")
barlist( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.axis = .8,srt = 0)
add_letter("b")
```

**Figure 2b.** Differential bacterial abundance across tissue and CVL microbiome week3 datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

\clearpage

```{r, echo = FALSE}
groupings <- lapply( names(datasets_all_samples)[-1] ,function(x){
  x <- read.csv2( paste0(PATH,"results/Sample_Louvain_clusters_",x,".csv"),row.names = 1 ) })
names(groupings) <- names(datasets_all_samples)[-1]

groupings_all_samples <- lapply(groupings,function(x){
  df <- data.frame(group=rep("NA",nrow(metadata)), row.names = rownames(metadata),stringsAsFactors = F)
  df[rownames(x),1] <- as.character(x[,1])
  return(df)
})
names(groupings_all_samples) <- names(groupings)

df <- as.data.frame(groupings_all_samples)
colnames(df) <- paste0( "louvain_",sub( "ASV_","",  sub( "_norma.*","", names(groupings_all_samples) )))
metadata <- cbind(metadata,df)
```

# Computing differential expression across microbiome datasets

```{r, echo = FALSE,fig.width=10,fig.height=9}
all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
datasets <- c( rep("tissue_v3",121) , rep("CVL_v3",121) , rep("CVL_v2",121) )
datasets <- paste0(datasets, "_",c(metadata$louvain_tissue_V3,metadata$louvain_CVL_V3,metadata$louvain_CVL_V2))

datasets <- factor( datasets[ colSums(all_microbiome)>0 ])
all_microbiome <- all_microbiome[,colSums(all_microbiome)>0]

NN <- min(table(datasets))
```

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(2,2)
group_means <- sapply( levels(datasets), function(x) rowMeans( all_microbiome[,datasets==x] ) )
cors <- cor(group_means)
h <- hclust( as.dist( (1-cors)/2 ),method = "ward.D2" )
o <- h$order

par(mar=c(6,6,2,2))
image(t(cors[o,o][nrow(cors):1,]),col=cor_pal,axes=F,border=NA,
      main="correlation matrix",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
text(  rep(par("usr")[c(1)]-.01*diff(par("usr")[c(1,2)]),length(h$labels)),
        rev((0:(length(h$labels)-1) )/(length(h$labels)-1)),
            labels = h$labels[o], srt = 0, adj = c(1,.5), xpd = TRUE, cex=1)
text(  (0:(length(h$labels)-1) )/(length(h$labels)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),length(h$labels)),
            labels = h$labels[o], srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
add_letter("a")

par(mar=c(6,4,2,2))
plot(as.dendrogram(h),las=1,ylab="height")
add_letter("b")
```

**Figure 3.** Comparrisson among patient groups across datasets. **(a)** Correlation matrix across sample groups. **(b)** Hierachical clustering of sample groups.

```{r, echo = FALSE,fig.width=10,fig.height=12}
res <- data.frame( matrix(0,nrow = 1,ncol = 6) )
for(i in levels(datasets)){
  for(j in rownames(all_microbiome) ){

    a <- all_microbiome[j,datasets == i]
    b <- all_microbiome[j,datasets != i]

    perc1 <- sum(all_microbiome[j,datasets == i]>0) / sum(datasets == i)
    perc2 <- sum(all_microbiome[j,datasets != i]>0) / sum(datasets != i)

    temp <- wilcox.test(x=a, y=b)

    fc <- log2( (mean(a)+1e-3) / (mean(b)+1e-3) )

    res <- rbind(res, setNames(c(j,i,fc,perc1,perc2,unlist(temp)[2] ),
                          c("bacteria","cluster","fc","perc.1","perc.2","pvalue")) )
    colnames(res) <- c("bacteria","cluster","fc","perc.1","perc.2","pvalue")
  }
}
res <- res[-1,]
res$pvalue <- as.numeric(res$pvalue)
res$perc.1 <- as.numeric(res$perc.1)
res$perc.2 <- as.numeric(res$perc.2)
res$perc.diff <- res$perc.1 - res$perc.2

res$fc <- as.numeric(res$fc)
res$FDR <- p.adjust(res$pvalue)
res <- res[order(res$pvalue),]
res <- res[abs(res$fc) > fct,]
res <- res[abs(res$pvalue) < pvt,]
res <- res[ (res$perc.1 > min_pct) | (res$perc.2 > min_pct) ,]
dim(res)

ord <- factor(sapply(unique(as.character(res$bacteria)),function(x){getcluster(all_microbiome, x, datasets)}))

mypar(1,1,mar=c(5,15,1,1))
plot_dots( all_microbiome , names(sort(ord)) , clustering = datasets,show_grid = T,cex.main=1,font.main=1,
          cex.row = .7,cex.col = .8,srt = 30)
add_letter("a")
```

**Figure 4.** Differential bacterial abundance across all groups and all microbiome datasets. The results are shown both as **a)** Dot plots and **b)** barplots Bacteria with log2FC above `r fct` and p-value below `r pvt` were considered significant and were sorted by the highest expression.

```{r, eval=F, echo = FALSE,fig.width=10, fig.height=10}
sample_order <- order(df$louvain_CVL_V3)

# zscore_datasets_all_samples <- lapply(datasets_all_samples,function(x){
#   zscore <- t(apply(x,1,function(i){scale(i,T,T)}))
#   zscore[ zscore > 6 ] <- 6
#   zscore[ zscore < -6 ] <- -6
#   rownames(zscore) <- rownames(x); colnames(zscore) <- colnames(x)
#   zscore[is.nan(zscore)] <- -6
#   return(zscore)
# })

mypar(3,1,mar=c(1,2,1,0.5))

image(t(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_tissue_V3)), pch=16,
        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_tissue_V3)), pch=16,
        col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)

image(t(datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V3)), pch=16,
        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V3)), pch=16,
        col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)

image(t(datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
        [ names(sort(ord)),sample_order]),
      col=heat_pal,axes=F,border=NA,
      main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
        rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V2)), pch=16,
        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
        rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
            length(df$louvain_CVL_V2)), pch=16,
        col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)

# **Figure 5.** Comparisson of microbiome datasets, showing only the significant bacteria. Samples are ordered by the CVL3 groupings. The colors represent their respective bacterial groupings for each dataset.

```

\clearpage

```{r, echo = FALSE,fig.width=8,fig.height=8}
metadata_parameters <- c("Antibiotics","Antibiotics_v2","Antibiotics_v3",
                         "Bacterio_KOH_v2","Bacterio_KOH_v3","Bacterio_TV_v3",
                         "BV_Clue_Cells_v2","BV_Clue_Cells_v3",
                         "BV_Diagnosis_v2","BV_Diagnosis_v3",
                         "BV_Lactobacillus_v2","BV_Lactobacillus_v3",
                         "BV_Monbilicus_v2","BV_Monbilicus_v3",
                         "BV_Vaginal_Garda_v2","BV_Vaginal_Garda_v3",
                         "blood_cervix_v2","blood_cervix_v3",
                         "cervical_inflammation_v2","cervical_inflammation_v3",
                         "HIVstatus","Contraception",
                         "CT_v2","CT_v3","regular_partner",
                         "vaginal_douching_v2","vaginal_douching_v3",
                         "louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3",
                         "group_M9_CVLv3","group_M6_CVLv3","group_M4_CVLv3","joint_clustering")

Ftest <- sapply(metadata_parameters,function(x){
  temp <- sapply(metadata_parameters,x=x,function(x,y){
    if (x==y){
      return( 1 )
    } else {
    f1 <- as.character(metadata[,x])
    f1[is.na(f1)] <- "NA"
    f1[f1==""] <- "NA"

    f2 <- as.character(metadata[,y])
    f2[is.na(f2)] <- "NA"
    f2[f2==""] <- "NA"

    f11 <- f1 [ (f1!="NA") & (f2!="NA") ]
    f22 <- f2 [ (f1!="NA") & (f2!="NA") ]
    contT1  <- table(f11,f22)
    print(x)
    print(y)
    return( fisher.test(contT1,simulate.p.value = T)$p.value )
    # return( chisq.test(contT1,simulate.p.value = T)$p.value )
    }
  })
  return(temp)
})

clean_Ftest <- Ftest
# clean_Ftest <- ( clean_Ftest + t(clean_Ftest) ) / 2
clean_Ftest <- -log10(clean_Ftest)
clean_Ftest[clean_Ftest > 5] <- 5
# clean_Ftest[ clean_Ftest > 0.05 ] <- 1

mypar(1,1,mar=c(12,1,1,12))
image( t(clean_Ftest [nrow(Ftest):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(Ftest)) , 
       labels = rownames(Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(Ftest)) , par("usr")[3]-.02 , 
      labels = rownames(Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
```

**Figure 6.** Association analysis across several patient categorical metadata parameters, including patient groupings annotations from microbiome.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(6,4,mar=c(1,4,1,4) )
x <- as.character(metadata[,"Contraception"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
plot_sankey( data.frame(x1,y1),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "Contraception" , "Antibiotics" ),pos=3,xpd=T)

x <- as.character(metadata[,"BV_Monbilicus_v2"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
plot_sankey( data.frame(x1,y1),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "BV_Monbilicus_v2" , "Antibiotics" ),pos=3,xpd=T)

x <- as.character(metadata[,"HIVstatus"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Antibiotics"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
plot_sankey( data.frame(x1,y1),color_by = 2,
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(10))
text(c(0,1),c(1,1),labels = c( "HIVstatus" , "Antibiotics_v2" ),pos=3,xpd=T)

x <- as.character(metadata[,"blood_cervix_v3"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"cervical_inflammation_v3"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
plot_sankey( data.frame(x1,y1),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(9))
text(c(0,1),c(1,1),labels = c( "blood_cervix_v3" , "cervical_inflammation_v3" ),pos=3,xpd=T)

```

**Figure 7.** A few examples of significant association between metadata parameters shown as sankey plots.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}
mypar(min(length(metadata_parameters),6),4,mar=c(1,4,1,4))

# metadata$Contraception
#Antibiotics usage
for(i in metadata_parameters){
  empty_plot()
  text( par("usr")[2] , mean(par("usr")[3:4]),labels = sub("_","\n",i),pos=2 )
  for(j in c("louvain_tissue_V3","louvain_CVL_V2","louvain_CVL_V3")){
    x <- as.character(metadata[,i])
    x[is.na(x)] <- "NA"
    y <- as.character(metadata[,j])
    y[is.na(y)] <- "NA"
    x1 <- x [ (x!="NA") & (y!="NA") ]
    y1 <- y [ (x!="NA") & (y!="NA") ]
    plot_sankey( data.frame(x1,y1),
             plot_labels = T, plot_weights = F,order_1_by = "total",
             order_2_by = "total",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal=hue_pal()(13))
    text(c(0,1),c(1,1),labels = c(i,j),pos=3,xpd=T)
  }
}
```

**Figure 8.** Sankey plots for all tested associations between the patient groups identified in in the microbiome datasets.

\clearpage

```{r, eval=F, echo = FALSE,fig.width=10,fig.height=13}
myletter <- letters

for(i in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected")){

  cat(paste0("Correlation between RNAseq and ",sub("_norm.*","",i)))

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- datasets_all_samples[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
  TRX <- TRX[ top_vars_TRX , ]
  dim(TRX)

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  # hist(cors,xlim=c(-1,1),breaks = 100)
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- bacteria_use

  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
  # mypar(mar=c(4,15,2,2))
  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  NESes <- NESes[rowSums(NESes!=0) >= 10 , colSums(NESes!=0) >= 10]

  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
  o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order

  NESes <- NESes[o_bacs,o_kegg]

  mypar(1,1,mar=c(12,2,2,12))
  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
  add_letter(myletter); myletter <- myletter[-1]
  # empty_plot()
  cat("\\clearpage")

}

# **Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown. Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.

```

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=10}

all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])

cors <- cor(t(all_microbiome[rowSums(all_microbiome > 0)>=10,]))

#Graph construction  
k <- 5 # k for KNN
P <- 0.2 # prunning threshold of SNN

knn <- RANN::nn2(cors,k = k)
# knn <- RANN::nn2(t(zscore),k = k)
nn <- matrix(0, nrow(knn$nn.idx), nrow(knn$nn.idx))
for(idx in 1:nrow(knn$nn.idx) ){  nn[idx,knn$nn.idx[idx,]] <- 1 }

#Compute SNN graph from KNN
jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))

#Prune SNN graph based on the Jaccard index
pSNN <- jSNN
pSNN <- pSNN / max(pSNN)

# hist(pSNN[pSNN>0],breaks = seq(0,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="SNN jaccard",xlim=c(0,1))
# abline(v=P,col="red",lty=2,xpd=F)
# 
# hist(cors[pSNN>0],breaks = seq(-1,1,length.out = 50),
#      border=NA,col="grey60",las=1,
#      xaxs="i",yaxs="i",
#      main="correlation coeficient",xlim=c(-1,1))
# abline(v=0,col="red",lty=1,xpd=F)

pSNN[ pSNN < P ] <- 0
pSNN[ pSNN > P ] <- cors[pSNN > P]
pSNN[ pSNN < 0 ] <- 0

g <- graph_from_adjacency_matrix(pSNN,mode = "undirected",diag = F,weighted = T)
```

# Visualise the data

```{r, echo = FALSE,fig.width=10,fig.height=9}
mypar(2,2,mar=c(0,0,2,10))

set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=1)

# cl <- igraph::cluster_louvain(g)
cl <- igraph::walktrap.community(g)

table(cl$membership)

plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(cl$membership)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=l,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(x = par("usr")[c(2)]+.2*diff(par("usr")[c(1,2)]),
       y = par("usr")[c(4)],
       legend = levels(factor(cl$membership)),
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
title("Louvain clusters",cex.main=1,font.main=1)
```

**Figure 9a.** Bacterial community embedding of `r k`-SNN graph clustered using walktrap community detection algorithm based on the normalized bacterial abundances across all microbiome datasets (including samples not present in all datasets).

```{r, echo = FALSE,fig.width  = 10, fig.height=12}
mypar(3,6,mar=c(1,1,2,1))
for(j in sort(unique(cl$membership)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(cors)[ cl$membership == j ],collapse = "\n"),adj=c(0,1),cex=.7)
  title(main=paste0("community",j))
}
```

**Figure 9b.** List of bacterial belongin to each community.

```{r, echo = FALSE,fig.width  = 10, fig.height=5}
taxa <- read.csv2("../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
rownames(cors) %in% rownames(taxa)

mypar(2,2,mar=c(1,10,2,10))
x <- taxa[match(rownames(cors),rownames(taxa)),2]
y <- taxa[match(rownames(cors),rownames(taxa)),3]
plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
       plot_labels = T, plot_weights = F,order_1_by = "total",
       order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
       pal=hue_pal()(length(unique(x))))

plot_sankey( data.frame( y[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
     plot_labels = T, plot_weights = F,order_1_by = "total",
     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
     pal=hue_pal()(length(unique(y))))

plot_sankey( data.frame( x[y!="NA"] , cl$membership[y!="NA"]), color_by = 1,
     plot_labels = T, plot_weights = F,order_1_by = "total",
     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
     pal=hue_pal()(length(unique(x))))

```

**Figure 9b.** Taxonomic annotation of bacterial communities.

```{r, eval=T, echo = FALSE,fig.width=10, fig.height=10}
sample_order <- order(df$louvain_CVL_V3)

module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
  return( rowsum( x[rownames(cors),] , group = cl$membership ) / c(table(cl$membership)) )
})

# mypar(3,1,mar=c(1,2,1,3))
# sample_order <- order(df$louvain_tissue_V3)
# image(t(module_means_per_dataset[["ASV_tissue_V3_normalized_batch_corrected"]]
#         [ ,sample_order]),
#       col=heat_pal,axes=F,border=NA,
#       main="tissue week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_tissue_V3)), pch=16,
#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
# points( (0:(length(df$louvain_tissue_V3)-1) )/(length(df$louvain_tissue_V3)-1),
#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_tissue_V3)), pch=16,
#         col=pal[factor(df$louvain_tissue_V3)][sample_order] ,xpd=T)
# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
# 
# 
# sample_order <- order(df$louvain_CVL_V3)
# image(t(module_means_per_dataset[["ASV_CVL_V3_normalized_batch_corrected"]]
#         [ ,sample_order]),
#       col=heat_pal,axes=F,border=NA,
#       main="CVL week 3",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_CVL_V3)), pch=16,
#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
# points( (0:(length(df$louvain_CVL_V3)-1) )/(length(df$louvain_CVL_V3)-1),
#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_CVL_V3)), pch=16,
#         col=pal[factor(df$louvain_CVL_V3)][sample_order] ,xpd=T)
# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
# 
# 
# sample_order <- order(df$louvain_CVL_V2)
# image(t(module_means_per_dataset[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]
#         [ ,sample_order]),
#       col=heat_pal,axes=F,border=NA,
#       main="CVL week 2",xlab=" ",ylab="",line=.4,cex.main=1,font.main=1)
# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
#         rep(par("usr")[c(3)]-.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_CVL_V2)), pch=16,
#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
# points( (0:(length(df$louvain_CVL_V2)-1) )/(length(df$louvain_CVL_V2)-1),
#         rep(par("usr")[c(4)]+.01*diff(par("usr")[c(3,4)]),
#             length(df$louvain_CVL_V2)), pch=16,
#         col=pal[factor(df$louvain_CVL_V2)][sample_order] ,xpd=T)
# text(  par("usr")[c(2)] , seq(1,0,length.out = nrow(NESes)),
#         labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

mypar(3,3,mar=c(3,5,2,2))

for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
          "ASV_CVL_V3_normalized_batch_corrected",
          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
    temp_group <- as.character(df[,i])
    temp_group[ is.na(temp_group) ] <- "NA"
    temp_group[ temp_group == "" ] <- "NA"
    temp_group <- factor(temp_group)

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}

```

```{r, eval=T, echo = FALSE,fig.width=10, fig.height=14}
mypar(7,3,mar=c(3,5,2,2))

for( i in c("louvain_tissue_V3","louvain_CVL_V3","louvain_CVL_V2")){
  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
          "ASV_CVL_V3_normalized_batch_corrected",
          "ASV_CVL_V2_normalized_NOT_batch_corrected")){
    temp_group <- as.character(df[,i])
    temp_group[ is.na(temp_group) ] <- "NA"
    temp_group[ temp_group == "" ] <- "NA"
    temp_group <- factor(temp_group)

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}

for( i in c("group_M4_CVLv3","group_M6_CVLv3","group_M9_CVLv3","joint_clustering")){
  for( j in c("ASV_tissue_V3_normalized_batch_corrected",
          "ASV_CVL_V3_normalized_batch_corrected",
          "ASV_CVL_V2_normalized_NOT_batch_corrected")){

    temp_group <- as.character(metadata[,i])
    temp_group[ is.na(temp_group) ] <- "NA"
    temp_group[ temp_group == "" ] <- "NA"
    temp_group <- factor(temp_group)

    barlist(data=module_means_per_dataset[[j]],
            genes=rownames(module_means_per_dataset[[j]]),
            labels = paste0("bac.com.",rownames(module_means_per_dataset[[j]])),
            clustering=temp_group, main=paste0(i,"\n", sub("_norm.*","",j)),srt=45)
  }
}
```

**Figure 5.** Comparisson of microbiome communities across all datasets and all clustering methods.

```{r, echo = FALSE,fig.width=10,fig.height=8}
myletter <- letters

NESse_list <- list()

for(i in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected")){

  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- module_means_per_dataset[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
  TRX <- TRX[ top_vars_TRX , ]
  dim(TRX)

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  # hist(cors,xlim=c(-1,1),breaks = 100)
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
  # mypar(mar=c(4,15,2,2))
  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]

  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
  # NESes <- NESes[o_bacs,o_kegg]

  NESse_list[[i]] <- NESes

  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
  # add_letter(myletter); myletter <- myletter[-1]
  # empty_plot()
  # cat("\\clearpage")

}

common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[colSums(x!=0)>=1] ) ) )

layout(matrix(c(1,2,3,4,4),nrow = 1))
par(mar=c(6,1,2,0))
for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(1,0,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

cat("\\clearpage")

#Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

#Combined representation of the two datasets
layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
par(mar=c(6,30,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
non_zeros <- colSums(TRX)!=0
TRX <- TRX [ top_vars_TRX , non_zeros ]
pathway_means <- t(sapply( colnames(x) , function(i){
  i <- gmt[[i]]
  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
  if(ncol(i)>1){
    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
    i <- colMeans(i)
  } else {
    i <- (i - min(i)) / (max(i) - min(i))
  }
  return( i - min(i) )
} ))
plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
        genes=rownames(pathway_means),
        clustering=metadata$joint_clustering[non_zeros],
        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))

par(mar=c(6,2,3,2))
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
# empty_plot(xaxs="i",yaxs="i",
#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

```

**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the Hallmark MSigDB gene annotation database. This, in turn, will result in a matrix associating every bacteria with every Hallmark MSigDB process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown.

\clearpage

```{r, echo = FALSE,fig.width=10,fig.height=8}
myletter <- letters

NESse_list <- list()

for(i in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected")){

  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))

  TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
  bac_dataset <- module_means_per_dataset[[i]]
  dge <- read.csv2(paste0(PATH,"/results/DGE_ASV_CVL_V2_normalized_NOT_batch_corrected.csv"),row.names = 1)
  samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i),] > 0)==2 ]

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
  top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
  # plot(apply(TRX,1,mean),apply(TRX,1,var),col=ifelse(rownames(TRX) %in% top_vars_TRX,"red","grey20"),cex=.2)
  TRX <- TRX[ top_vars_TRX , ]
  dim(TRX)

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
  top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  # hist(cors,xlim=c(-1,1),breaks = 100)
  dim(cors)

  bacteria_use <- colnames(cors)
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]
  # mypar(mar=c(4,15,2,2))
  # boxplot( t(pvalues[rev(top_pathways),]),cex=.1,horizontal=T,las=1 )

  # pvalues[ pvalues < -log10(0.05)] <- 0
  # sort(rowSums(pvalues>0),T)
  # sort(colSums(pvalues>0),T)
  NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  NESes[ pvalues < -log10(0.05) ] <- 0
  # NESes <- NESes[rowSums(NESes!=0) >= 2 , colSums(NESes!=0) >= 2]

  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes)))/2 ),"ward.D2")$order
  # o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
  # o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
  # NESes <- NESes[o_bacs,o_kegg]

  NESse_list[[i]] <- NESes

  # image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
  #        axes=F,border=NA,main=i,xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
  #       labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
  # text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
  #       labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)
  # add_letter(myletter); myletter <- myletter[-1]
  # empty_plot()
  # cat("\\clearpage")

}

common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[ colSums(x!=0)>=2 ] ) ) )

layout(matrix(c(1,2,3,3,3),nrow = 1))
par(mar=c(6,1,3,0))

for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="KEGG")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

cat("\\clearpage")

#Combined representation of the two datasets
layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
par(mar=c(6,30,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
non_zeros <- colSums(TRX)!=0
TRX <- TRX [ top_vars_TRX , non_zeros ]
pathway_means <- t(sapply( colnames(x) , function(i){
  i <- gmt[[i]]
  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
  if(ncol(i)>1){
    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
    i <- colMeans(i)
  } else {
    i <- (i - min(i)) / (max(i) - min(i))
  }
  return( i - min(i) )
} ))
plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
        genes=rownames(pathway_means),
        clustering=metadata$joint_clustering[non_zeros],
        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))

par(mar=c(6,2,3,2))
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
# empty_plot(xaxs="i",yaxs="i",
#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

```

**Figure 9.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown. Bacteria and pathways significant in less that 10 pathways and bacteria, respectivelly, were omitted.

\clearpage