### 2021-02-15 reports/rmarkdown/manuscript/manuscript_template.Rmd
fixes plots and adds picrust

---
title: "Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(out.width = '100%',
                      dpi=600,
                      fig.width=10,
                      echo = FALSE,
                      # fig.pos = 'p', # Places figures on their own pages
                      eval = TRUE, 
                      results = 'hide',
                      warning = FALSE,
                      message = FALSE )

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
library(igraph)
library(dplyr)
library(stringr)
library(rafalib)
library(scales)
# remotes::install_github("czarnewski/niceRplots",force=T)
library(niceRplots)

#color palettes
pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

Abstract
===============================================================================

Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract 
test test test

\clearpage

Introduction
================================================================================

Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction ([@Johnston1972]), Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction Introduction **Introduction Introduction** *Introduction* Introduction ([@Darwin1859]^,^[@Bumpus1898]) .

Problem / question to answer

\clearpage

Results
===============================================================================

```{r, echo = FALSE, eval = TRUE}
# Load reasult from the analysis
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]

SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
clean_Ftest <- read.csv2("../../../results/metadata_association_results.csv",row.names = 1)
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")

picrust_data <- read.table("../../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
picrust_dge <- read.csv("../../../data/Picrust2_top_functions.csv",header = T,row.names = 1)
```

```{r, echo = FALSE, eval = TRUE}
#Assign names to patient groups
big_groups <- c(
  L.inners=    "5",
  L.inners2=   "4",
  L.crispatus= "2",
  Gard=        "3",
  Gard.Prev=   "6",
  Gard.Prev2=  "8",
  Gard.Lacto=  "7",
  Patho=       "1"
)

group_annotation <- as.character(patient_groups)
group_annotation <- names(big_groups)[match(group_annotation,big_groups)]
group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))

metadata$group_annotation <- group_annotation
```

```{r, echo = FALSE, eval = TRUE}
bac_communities <- factor(paste0("BC",sprintf("%02d",bac_communities)))
bacs <- names(bac_communities)

#Assign names to bacterial communities
BC_groups <- c(
  BC01= "BC01_Bifidobacterium",
  BC02= "BC02_Mycoplasma",
  BC03= "BC03_Lactobacillus.spp",
  BC04= "BC04_Escherichia.sp",
  BC05= "BC05_Other1",
  BC06= "BC06_Other2",
  BC07= "BC07_Corynebacterium.spp",
  BC08= "BC08_Other3",
  BC09= "BC09_Ureaplasma.sp",
  BC10= "BC10_Campylobacter",
  BC11= "BC11_Mobiluncus_BVAB",
  BC12= "BC12_Faecalibacterium",
  BC13= "BC13_Dialister",
  BC14= "BC14_P.timonesis",
  BC15= "BC15_T.scotoductus"
)

bac_communities <- as.character(bac_communities)
bac_communities <- BC_groups[match(bac_communities,names(BC_groups))]
bac_communities <- as.factor(setNames(bac_communities,bacs))
```

**Joint analysis of vaginal microbiome reveals distinct patient subgroups**

To understand the longitudinal and tissue-specific microbiome profile in vaginal samples, `r nrow(metadata)` adult female sex workers were enrolled in [...]. Among those, `r sum(metadata$HIVstatus=="pos")` were previously tested positive for HIV during the cohort's sampling procedure. [Describe here what was done and when, which samples, which tissues].

To be able to better undertand the differences in microbiome profile across all datasets collected, we  performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see "Methods" section for details). A total of `r length(unique(bac_communities))` bacterial communities were identified.

Noticebly, bacterial community `r bac_communities["Lactobacillus iners"]` consisted only of Lactobacillus species (*`r names(bac_communities)[bac_communities %in% bac_communities["Lactobacillus iners"]]`*).

Patients were thus subdivided into `r length(unique(group_annotation))` groups,

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

\clearpage

**Indentification of bacterial communities metabolic processes linked to Lactobacilli**

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

\clearpage

**Indentification of bacterial communities metabolic processes linked to Lactobacilli**

Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results Results

\clearpage

Discussion
================================================================================

I have analysed data collected by Herman Bumpus [@Bumpus1898] on the relationship between sparrow (*Passer domesticus*) total length and surival following an unusually severe storm. I found that sparrows that died in the storm were longer than sparrows that survived, which suggests that higher sparrow body length decreased survival. Of course, it is not possible to definitively conclude a causal relationship between any aspect of body size and sparrow survival<!--- BD Note: maybe explain this better --->, and even the available data collected by Bumpus would permit a more thoughtful analysis than that conducted in this study (see [Appendix Table 1](#appendix)).

<!---

Here is one way to add some more detailed comments into the manuscript itself, though this can also be done within the text (see above).

--->

Overall, this document demonstrates how high quality, professional looking documents can be written using Rmarkdown. The [underlying code](https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/ms.Rmd) for this manuscript is publicly available, along with [accompanying notes](https://stirlingcodingclub.github.io/Manuscripts_in_Rmarkdown/Rmarkdown_notes.html) to understand how it was written. By using Rmarkdown to write manuscripts, authors can more easily use version control (e.g., git) throughout the writing process. The ability to easily integrate citations though BibTeX, LaTeX tools, and dynamic R code can also make writing much more efficient and more enjoyable. Further, obtaining the benefits of using Rmarkdown does not need to come with the cost of isolating colleagues who prefer to work with Word or LaTeX because Rmarkdown can easily be converted to these formats (in the case of Word, with the push of a button). By learning all of the tools used in this manuscript, readers should have all of the necessary knowledge to get started writing and collaborating in Rmarkdown.

\clearpage

Methods
================================================================================

\clearpage

References
================================================================================

<div id="refs"></div>

\clearpage

<a name="appendix">Appendix Table 1</a>
================================================================================

An example table is shown below, which includes all of the variables collected by @Bumpus1898 for the first 10 measured sparrows. The full data set can be found online in [GitHub](https://github.com/StirlingCodingClub/Manuscripts_in_Rmarkdown/blob/master/data/Bumpus_data.csv).

```{r, echo = FALSE, fig.pos='H',fig.show='hold'}
library(knitr)
```

\clearpage

FIGURES (MAIN)
================================================================================

## Figure 1

```{r, echo = FALSE, eval = TRUE , fig.asp=9/10}
# Define Layout and set labels
layout(matrix(c(1,1,1,1,      3,3,3,3,          3,4,4,4,
                2,2,2,2,      3,3,3,3,          3,5,5,5,
                6,6,6,6,      9,9,9,9,          9,9,9,9,
                7,7,7,7,     10,10,10,10,     10,10,10,10,
                8,8,8,8,     10,10,10,10,     10,10,10,10 ) , ncol= 12,byrow = T),
       widths = 1,heights = c(1,1,.7,.7,.7))
# layout.show(12)
figlabels <- letters

#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0,0,0,0))
fig1a <- magick::image_read("schematics.pdf",density = 900)
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

####################
# PATIENT GROUPING #
####################
g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

par(mar=c(0,0,0,5))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10,main="", 
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend("topright",title.adj = 0,
       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

##############################
# METADATA CROSS-ASSOCIATION #
##############################
par(mar=c(12,1,1,12))
image( t(clean_Ftest [nrow(clean_Ftest):1,]),axes=F,
       col=colorRampPalette(c("grey95","grey70","orange3","firebrick"))(90) )
text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(clean_Ftest)) ,
       labels = rownames(clean_Ftest), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
text( seq(0,1,length.out = nrow(clean_Ftest)) , par("usr")[3]-.02 ,
      labels = rownames(clean_Ftest), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,1,1,1))
empty_plot(frame = T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

empty_plot(frame = T)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,2,1))
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
xL <- rowsum(x , grepl("Lactobacillus",rownames(x)) )[2,] / sum(grepl("Lactobacillus",rownames(x)))
xM <- rowsum(x , grepl("Mobiluncus",rownames(x)) )[2,] / sum(grepl("Mobiluncus",rownames(x)))
xG <- rowsum(x , grepl("Gardnerella",rownames(x)) )[2,] / sum(grepl("Gardnerella",rownames(x)))

mL <- metadata$BV_Lactobacillus_v3 ; mL[is.na(mL)] <- 0
mM <- metadata$BV_Monbilicus_v3    ; mM[is.na(mM)] <- 0
mG <- metadata$BV_Vaginal_Garda_v3 ; mG[is.na(mG)] <- 0

barlist( data = rbind(  BV=mL ,
                        counts_16S=xL),
         main = "Lactobacillus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

barlist( data = rbind(  BV=mM ,
                        counts_16S=xM),
         main = "Monbilicus",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

barlist( data = rbind(  BV=mG ,
                        counts_16S=xG),
         main = "Gardanerella",
         genes = c("BV","counts_16S"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(1,5,1,12))

shann <- vegan::diversity(t(x),index = c("shannon") )
simp <- vegan::diversity(t(x),index = c("simpson") )
invsimp <- vegan::diversity(t(x),index = c("invsimpson") )
divers <- as.matrix(rbind(shann,simp,invsimp) )
divers[is.infinite(divers)] <- 0
rownames(divers) <- c("Shannon","simpson","Inv Simpson")
barlist( data = divers,
         main = "",
         genes = c("Shannon","simpson","Inv Simpson"),
         clustering = group_annotation,
         draw_mean_lines=F,col = pal)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

#######################
# BACTERIAL ABUNDANCE #
#######################
par(mar=c(4,5,1,12))
temp <- t(t(2^x-1)/colSums(2^x-1))
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
temp[is.nan(temp)] <- 0
o <- order( rowSums(temp) , decreasing = T)
sample_o <- order( group_annotation )
temp <- temp[o,sample_o]*100
temp <- temp[-grep("NA",rownames(temp)),]
temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
legend(x = par("usr")[c(2)],
       y = par("usr")[c(4)],title.adj = 0,
       legend = rownames(temp),xjust = 0,yjust = 1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
       pt.bg = taxa_pal)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]
```

**Figure 1. Identification of patient groups.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

\clearpage

## Figure 2

```{r, echo = FALSE, eval = TRUE}
# Define Layout and set labels
layout(matrix(c(2 ,3,4,5,      6,7,17,17,          19,19,19,19,
                1,1,1,1,      1,8,17,17,          20,20,20,20,
                1,1,1,1,      1,9,18,18,         21,21,24,24,
                1,1,1,1,      1,10,18,18,         22,22,24,24,
                16,15,14,13,  12,11,23,23,         23,23,24,24 ) , ncol= 12,byrow = T),
       widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
# layout.show(12)
figlabels <- letters

#########################
# BACTERIAL COMMUNITIES #
#########################
gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
set.seed(1)
lB <- layout_with_fr(gB,niter=3000,start.temp=30)

par(mar=c(0,0,0,5))
plot( gB , vertex.label.cex=0.000001 , vertex.color = pal[factor(bac_communities)] ,
      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
      vertex.size=10,main="", 
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] )
legend("topright",title.adj = 0,
       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

par(mar=c(1,1,1,1))
for(j in sort(unique(bac_communities)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
  title(main=paste0("community",j),cex.main=.7)
}
for(i in 1:(18-length(unique(bac_communities)))){empty_plot()}

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"

par(mar=c(1,7,2,3))
x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
#        plot_labels = T, plot_weights = F,order_1_by = "total",
#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#        pal=hue_pal()(length(unique(x))))

# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
#      plot_labels = T, plot_weights = F,order_1_by = "total",
#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#      pal=hue_pal()(length(unique(y))))

plot_sankey( data.frame( x[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
     plot_labels = T, plot_weights = F,order_1_by = "total",
     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
     pal=hue_pal()(length(unique(x))))

add_letter(figlabels[1]); figlabels <- figlabels[-1]

##############################################
# BACTERIAL COMMUNITY DIFFERENTIAL ABUNDANCE #
##############################################
par(mar=c(1,7,1,3))

module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
  return( rowsum( x[rownames(SNN_bacteria),] , group = bac_communities ) / c(table(bac_communities)) )
})

for( j in c("ASV_tissue_V3_normalized_batch_corrected",
        "ASV_CVL_V3_normalized_batch_corrected",
        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
  temp_group <- as.character(group_annotation)
  temp_group[ is.na(temp_group) ] <- "NA"
  temp_group[ temp_group == "" ] <- "NA"
  temp_group <- factor(temp_group)

  barlist(data=module_means_per_dataset[[j]],
          genes=rownames(module_means_per_dataset[[j]]),col = pal,
          clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
}
add_letter(figlabels[1]); figlabels <- figlabels[-1]

empty_plot()

#####################
# BACTERIAL PICRUST #
#####################
par(mar=c(1,7,1,3))
# 
# library(keggorthology)
# data("KOgraph")
# annot <- as.data.frame(unlist(nodeData(KOgraph,attr="tag")))
# colnames(annot)[1] <- "KEGGID"
# annot$pathway <- rownames(annot)
# annot$KEGGID <- paste0("K",annot$KEGGID)
# 
# library(KEGGREST)

download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/ko_info.tsv.gz","../../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/pathway_mapfiles/KEGG_pathways_to_KO.tsv","../../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/KEGG_pathways_info.tsv.gz","../../../data/KEGG_pathways_info.tsv.gz")
system("gunzip ../../../data/*.gz")

kegg_pathway_annot <- fgsea::gmtPathways('../../../data/KEGG_pathways_to_KO.tsv')

kegg_names <- read.delim('../../../data/KEGG_pathways_info.tsv',header = F)

ko_names <- readLines('../../../data/ko_info.tsv')
ko_names <- strsplit(ko_names,"\t")

path_trend <- t(sapply(kegg_pathway_annot,function(x){
  x <- rownames(picrust_data)[rownames(picrust_data) %in% x]
  if(length(x)>=2){
    return(colSums(picrust_data [ x , ]))
  } else {
    return(rep(0, ncol(picrust_data)))
  }
}))

path_trend <- path_trend[rowSums(path_trend)!=0 , ]
rownames(path_trend) <- kegg_names[ match(rownames(path_trend),kegg_names[,1]),2]

library(edgeR)
design <- model.matrix(~  group_annotation, data=metadata)

y <- DGEList(counts=floor(path_trend), remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

lrt <- glmLRT(fit,coef=1:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(logFC.Gard=0,top)

top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

norm_path_trend <- log2( t( t(path_trend) / colSums(path_trend) ) * 1000 + 1 )
norm_path_trend <- norm_path_trend[top_dge, ]
norm_path_trend <- t(apply(norm_path_trend,1,function(x){scale(x,T,T)}))
norm_path_trend[norm_path_trend > 5] <- 5
colnames(norm_path_trend) <- colnames(path_trend)
sample_use <- colnames(norm_path_trend)

mypar(2,4,mar=c(1,.2,2,.2))
h <- hclust( as.dist( (1- cor(t(norm_path_trend)))/2 ), method = "ward.D2")
plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
abline(v=1,xpd=F,col="red",lty=2)
gene_module <- cutree(h, h = 1)

points( rep(0,length(gene_module)),
        seq(length(gene_module),1,length.out = length(gene_module)),
        col=taxa_pal[factor(gene_module[h$order])],
        pch=16,cex=.5)

image( t(norm_path_trend[h$order,][nrow(norm_path_trend):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16)

par(mar=c(4,4,2,2))
set.seed(1)
UMAP_picrust <- uwot::umap(t(norm_path_trend),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[group_annotation[sample_use]],main="UMAP on top functional profile",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
```

**Figure 2. Identification and characterization of vaginal bacterial communities.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

\clearpage

## Figure 3

```{r, echo = FALSE, eval = TRUE}
# Define Layout and set labels
layout(matrix(c(1) , ncol= 1,byrow = T),
       heights = c(1))
figlabels <- letters

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
TRX_counts <- TRX_counts[,sample_use]

library(edgeR)

design <- model.matrix(~  group_annotation, data=metadata[sample_use,])

y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

lrt <- glmLRT(fit,coef=2:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(logFC.Gard=0,top)

top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

top_TRX_counts <- cpm(y)[top_dge, ]
top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
top_TRX_counts[top_TRX_counts > 5] <- 5
colnames(top_TRX_counts) <- colnames(TRX_counts)

mypar(2,4,mar=c(1,.2,2,.2))
h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
plot( rev(as.dendrogram(h)) , horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=F)
abline(v=2,xpd=F,col="red",lty=2)
gene_module <- cutree(h, h = 2)

points( rep(0,length(gene_module)),
        seq(length(gene_module),1,length.out = length(gene_module)),
        col=taxa_pal[factor(gene_module[h$order])],
        pch=16,cex=.5)

image( t(top_TRX_counts[h$order,][nrow(top_TRX_counts):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16)

set.seed(1)
UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
                 metric = "correlation",min_dist = 0.1,spread = 5,
                 negative_sample_rate = 10)
plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="UMAP on top DGE RNAseq",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
```

**Figure 1. Identification of patient groups.**
(**a**) Schematic representation of ########## .
(**b**) Schematic representation of ########## .
(**c**) Schematic representation of ########## .
(**d**) Schematic representation of ########## .

\clearpage

FIGURES (SUPPL)
================================================================================

## Figure S1

```{r, echo = FALSE, eval = TRUE, fig.asp=1}
figlabels <- letters
temp_legend <- NULL

layout(matrix(c(1,7,
                2,7,
                3,7,
                4,8,
                5,8,
                6,8) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))

for(i in c("ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected",
           "ASV_tissue_V3_normalized_batch_corrected")){
  x <- datasets_all_samples[[i]]

  par(mar=c(1,5,1,1))
  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
  barlist( data = divers[,order(group_annotation)],
         main = "",
         genes = c("Shannon","simpson","Inv Simpson"),
         # clustering = group_annotation,
         draw_mean_lines=F,col = pal[factor(group_annotation[order(group_annotation)])])

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(3,5,1,1))
  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
  # taxa[is.na(taxa)] <- "NA"
  # taxa <- taxa[rownames(temp),]
  # temp <- rowsum(temp , taxa$Class)
  temp[is.nan(temp)] <- 0

  if(is.null(temp_legend)){ 
    temp_legend <- temp   #Use the same legend from the 1st dataset
  }

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  # temp <- temp[-grep("NA",rownames(temp)),]
  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
  # legend(x = par("usr")[c(2)],
  #        y = par("usr")[c(4)],title.adj = 0,
  #        legend = rownames(temp),xjust = 0,yjust = 1,
  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
  #        pt.bg = taxa_pal)
}

par(mar=c(4,0,1,0))
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],title.adj = 0,
         legend = levels(group_annotation),xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
         pt.bg = pal)

empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],title.adj = 0,
         legend = rownames(temp),xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          title="top abundant bacteria",
         pt.bg = taxa_pal)

```

\clearpage

## Figure S2

```{r, echo = FALSE, eval = TRUE, fig.asp=.5}
figlabels <- letters
temp_legend <- NULL

mypar(1,4,mar=c(12,1,1,0))

for(i in c("ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected",
           "ASV_tissue_V3_normalized_batch_corrected")){

  continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
  continuous_metadata <- metadata[,continuous_parameters]
  # continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )

  metadata_cor <- cor(t(module_means_per_dataset[[i]]),continuous_metadata,use = "pairwise.complete.obs")
  # metadata_cor_test <- cor.test(t(module_means_per_dataset[[i]]),continuous_metadata)

  image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
         breaks = seq(-1,1,length.out = 90),main=sub("_nor.*","",i),font.main=1,
         col=colorRampPalette(c("navy","grey95","firebrick"))(89) )
  # text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(metadata_cor)) , 
  #        labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
  text( seq(0,1,length.out = ncol(metadata_cor)) , par("usr")[3]-.02 , 
        labels = colnames(metadata_cor), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]
}

par(mar=c(12,0,1,0))
image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
         breaks = seq(-1,1,length.out = 90),main="",
         col=colorRampPalette(c("white"))(89) )
text(  0 ,seq(1,0,length.out = nrow(metadata_cor)) , 
  labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
```

\clearpage

## Figure S3

```{r, echo = FALSE, eval = TRUE}
figlabels <- letters
mypar(1,2,mar=c(3,3,3,10))

datasets <- datasets_all_samples[grep("ASV",names(datasets_all_samples))]
datasets <- lapply(datasets,t)
datasets <- t( as.data.frame(datasets) )

group_means <- sapply(levels(group_annotation), function(x){
  rowMeans( datasets[ , group_annotation == x ] )
})

d <- as.dist( (1 - cor(group_means)) / 2 )
h <- hclust( d , method = "ward.D2")
plot( as.dendrogram(h) , horiz=T, xlab="height")
points( rep(0 , length(h$labels)) , 1:length(h$labels) , pch=21, cex=1.5,
        bg=pal[factor(h$labels[ h$order ],levels = h$labels)] )
add_letter(figlabels[1]); figlabels <- figlabels[-1]

datasets <- datasets_all_samples[grep("ASV",names(datasets_all_samples))]
datasets <- as.data.frame(datasets)

group_means <- sapply(levels(factor(bac_communities)), function(x){
  colMeans( datasets[ bac_communities == x , ] )
})

d <- as.dist( (1 - cor(group_means)) / 2 )
h <- hclust( d , method = "ward.D2")
plot( as.dendrogram(h) , horiz=T, xlab="height")
points( rep(0 , length(h$labels)) , 1:length(h$labels) , pch=21, cex=1.5,
        bg=taxa_pal[factor(h$labels[ h$order ],levels = h$labels)] )
add_letter(figlabels[1]); figlabels <- figlabels[-1]
```

\clearpage

## Figure S4

```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
figlabels <- letters
for(j in names(module_means_per_dataset)){
  mypar(2,8,mar=c(0.5,0.5,4,0.5))

  plot(UMAP_picrust,bg=pal[group_annotation],main="patient groups",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  for(i in 1:nrow(module_means_per_dataset[[j]])){

    xx <- module_means_per_dataset[[j]][i,]
    xx <- xx / max(xx)
    o <- order(xx)
    plot(UMAP_picrust[o,],
         col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
         main=rownames(module_means_per_dataset[[j]])[i],
         pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
  }
}
```

\clearpage

## Figure S5

```{r, echo = FALSE, eval = TRUE,fig.asp=.3}
figlabels <- letters
for(j in names(module_means_per_dataset)){
  mypar(2,8,mar=c(0.5,0.5,4,0.5))

  plot(UMAP_TRX,bg=pal[group_annotation[sample_use]],main="patient groups",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  for(i in 1:nrow(module_means_per_dataset[[j]])){

    xx <- module_means_per_dataset[[j]][i,][sample_use]
    xx <- xx / max(xx)
    o <- order(xx)
    plot(UMAP_TRX[o,],
         col=colorRampPalette( c("grey95","navy")) (99)[ 1 + (xx[o]*98) ],
         main=rownames(module_means_per_dataset[[j]])[i],
         pch=16,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F,font.main=1,cex.main=1)
  }
}
```

\clearpage

TABLES (MAIN)
================================================================================

## Table 1

```{r, echo = FALSE, eval = TRUE}
```

\clearpage

## Table 2

```{r, echo = FALSE, eval = TRUE}
```

\clearpage

## Table 3

```{r, echo = FALSE, eval = TRUE}
```

\clearpage

TABLES (SUPPL)
================================================================================

## Table S1

```{r, echo = FALSE, eval = TRUE}
```

\clearpage

## Table S2

```{r, echo = FALSE, eval = TRUE}
```

\clearpage

## Table S3

```{r, echo = FALSE, eval = TRUE}
```

\clearpage