### 2021-02-20 reports/rmarkdown/05_picrust_analysis.Rmd
adds picrust and RNAseq to manuscript

---
title: "Piecrust_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r }
picrust_data <- read.table("../../data/pred_metagenome_unstrat.tsv",header = T,row.names = 1)
picrust_dge <- read.csv("../../data/Picrust2_top_functions.csv",header = T,row.names = 1)
metadata <- read.csv2("../../results/metadata_integration.csv",row.names = 1)

download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/ko_info.tsv.gz","../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/pathway_mapfiles/KEGG_pathways_to_KO.tsv","../../data/KEGG_pathways_to_KO.tsv")
download.file("https://raw.githubusercontent.com/picrust/picrust2/master/picrust2/default_files/description_mapfiles/KEGG_pathways_info.tsv.gz","../../data/KEGG_pathways_info.tsv.gz")
system("gunzip ../../data/*.gz")

kegg_pathway_annot <- fgsea::gmtPathways('../../data/KEGG_pathways_to_KO.tsv')

kegg_names <- read.delim('../../data/KEGG_pathways_info.tsv',header = F)

ko_names <- readLines('../../data/ko_info.tsv')
ko_names <- strsplit(ko_names,"\t")
```

```{r }

#Sum counts to pathway level
path_trend <- t(sapply(kegg_pathway_annot,function(x){
  x <- rownames(picrust_data)[rownames(picrust_data) %in% x]
  if(length(x)>=2){
    return(colSums(picrust_data [ x , ]))
  } else {
    return(rep(0, ncol(picrust_data)))
  }
}))
path_trend <- path_trend[rowSums(path_trend)!=0 , ]
rownames(path_trend) <- kegg_names[ match(rownames(path_trend),kegg_names[,1]),2]

library(edgeR)

#GLM fit across all patient groups
metadata$joint_clustering <- factor(metadata$joint_clustering)
design <- model.matrix(~  joint_clustering, data=metadata)
y <- DGEList(counts=floor(path_trend), remove.zeros = T)
y <- calcNormFactors(y,method = "TMM")
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTagwiseDisp(y,design)
fit <- glmFit(y, design)

#Perform the test
lrt <- glmLRT(fit,coef=2:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("joint_clustering","",colnames(top))
write.csv2(top,"../../results/Picrust_DGE_table.csv",row.names = T)

#Filter significant bacteria
top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

#Log Normalize
norm_path_trend <- log2( t( t(path_trend) / colSums(path_trend) ) * 1000 + 1 )
norm_path_trend <- norm_path_trend[top_dge, ]

#Z-score transform
norm_path_trend <- t(apply(norm_path_trend,1,function(x){scale(x,T,T)}))
norm_path_trend[norm_path_trend > 5] <- 5
colnames(norm_path_trend) <- colnames(path_trend)

write.csv2(norm_path_trend,"../../results/Picrust_norm_path_trend.csv",row.names = T)
```