### 2021-08-16 reports/rmarkdown/manuscript/Figure1.Rmd

---
title: "Figure 1. Microbiome abundances"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))
#group_annotation <- as.character(metadata$Luminal_gr) %>% set_names()

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

#color palettes
#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")

pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
#pal <- c("#377EB8", "#4DAF4A", "#FF7F00", "#F781BF", "#984EA3", "#FFFF33", "#A65628",  "#999999")
#pal <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
#show_col(pal)
#show_col(taxa_pal)
```

```{r Figure 1, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=6 ,fig.asp=1, fig.pos="t"}
### A + B
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################

figlabels <- letters
temp_legend <- NULL

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                4,4,4,4,4,4,
                5,5,8,9,6,6,
                5,5,8,7,6,6,
                5,5,8,7,7,7) , ncol= 6,byrow = T),
                               widths = c(1.2,1,1.85,1.35,1,1.4),
                               heights = c(.7,1.5,.7,1.5,1,.8,.2))
# layout(matrix(c(1,1,1,1,1,5,
#                 2,2,2,2,2,5,
#                 3,3,3,3,3,6,
#                 4,4,4,4,4,6,
#                 7,7,8,8,8,6) , ncol= 6,byrow = T),
#                                widths = c(.5,1,1,1,1,1.1),
#                                heights = c(.5,1,.5,1,1.5))

#i <- ASV[2]
ASV <- list(c("Luminal_gr","ASV_CVL_V3_normalized_batch_corrected"),
         c("Tissue_gr","ASV_tissue_V3_normalized_batch_corrected")
          #"3"="ASV_CVL_V2_normalized_NOT_batch_corrected"
         )
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))
  #x <- x[,sample_use]
  #x <- datasets_all_samples[[2]]

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="", #bg="white", #cex=.6, 
         genes = c("Shannon","simpson","Inv Simpson"),
         # clustering = group_annotation,
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  if(i[2] == "ASV_CVL_V3_normalized_batch_corrected"){
    end <- table(pal[factor(group_annotation[order(group_annotation)])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
    #print(par("usr"))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1,5,1,1.5)) #b,l,t,r
  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
  # taxa[is.na(taxa)] <- "NA"
  # taxa <- taxa[rownames(temp),]
  # temp <- rowsum(temp , taxa$Class)
  temp[is.nan(temp)] <- 0

  if(is.null(temp_legend)){
    temp_legend <- temp   #Use the same legend from the 1st dataset
  }

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  # temp <- temp[-grep("NA",rownames(temp)),]
  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=gsub("_gr","",i[1]),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = lvl)] ,names.arg = rep("",ncol(temp)))

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)
  #text(x = c(2), y = 3, ']', srt = 90, cex = 3, xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T) # pos=4,

}
### C
############################
# TOTAL RELATIVE ABUNDANCE #
############################
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]
raw <- c("ASV_CVL_V3", "ASV_tissue_V3") %>% set_names()

raw_list <- raw %>%
  map(., ~read_csv(paste0("../../../results/",.x,"_raw_counts.csv")))  %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.5514392, "Other")) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_CVL_V3) %>%
    dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
    `row.names<-`(., NULL) %>% 
    column_to_rownames(var = "name") %>%
    as.matrix()

#colSums(r)

par(mar=c(1,5,2,0)) #b,l,t,r
barplot( r, border = NA ,ylab="abundance", xlab ="",
         col = taxa_pal[factor(rownames(r),levels = rownames(temp))])
title(main = "Total relative abundance", line = 0.5, cex.main = .9, font.main= 1)
mtext("Luminal        Tissue ", line = 0, cex=.6, side=1)
#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]
# legend(x = par("usr")[c(2)],
#          y = par("usr")[c(4)],title.adj = 0,
#          legend = rownames(r),xjust = 0,yjust = 1,
#          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
#           title="top abundant bacteria",
#          pt.bg = taxa_pal)

### D
#############################
# LUMINAL-TISSUE COMPARISON #
#############################
par(mar=c(0,4,3,2)) #b,l,t,r
x <- as.character(metadata[,"Luminal_gr"])
x[is.na(x)] <- "NA"
y <- as.character(metadata[,"Tissue_gr"])
y[is.na(y)] <- "NA"
x1 <- x [ (x!="NA") & (y!="NA") ]
y1 <- y [ (x!="NA") & (y!="NA") ]
df <- data.frame(x1,y1)
plot_sankey( df,color_by = 1, 
             plot_labels = T, plot_weights = F,order_1_by = "NULL",
             order_2_by = "disentangled",use_w1 = T,use_w2 = T,gap2v = 0,smoothing = .1,
             pal= pal)
text(c(0,1),c(1,1),labels = c( "Luminal" , "Tissue" ),pos=3, xpd=T)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

###########
# LEGENDS #
###########
par(mar=c(0,.5,0,0)) #b,l,t,r
empty_plot()
legend(x = -0.04, #par("usr")[c(1)],
       y = 1.04, #par("usr")[c(4)],title.adj = 0,
       #legend = sort(unique(group_annotation)),
       legend = c("L. crispatus/jensenii", "L. iners", "Gardnerella", "High diverse", "Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1, #xjust=1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "patient groups",
       pt.bg = pal)

par(mar=c(0,.4,1,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],#title.adj = 0,
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          #title="",
         pt.bg = taxa_pal)

par(mar=c(0,.4,1,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],#title.adj = 0,
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          #title="",
         pt.bg = taxa_pal[13:17])

```

```{r include=FALSE}
dev.copy2pdf(file=paste0("./images/","Figure1",".pdf"), 
    width = 6.5,
    height = 7 , paper = "a4" #units = "cm", res = 300 # a4 = 8.3 x 11.7 inches
)
```

```{r eval=FALSE, include=FALSE}
##########################
x <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
y <- datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]]
z <- datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]]

 # par(mar=c(1,5,1,1))
tempx <- t(t(2^x-1)/colSums(2^x-1))
tempy <- t(t(2^y-1)/colSums(2^y-1))
tempz <- t(t(2^z-1)/colSums(2^z-1))

tempx[is.nan(tempx)] <- 0
tempy[is.nan(tempy)] <- 0
tempz[is.nan(tempz)] <- 0
 # if(is.null(temp_legend)){
 #   temp_legend <- temp   #Use the same legend from the 1st dataset
#  }

temp1 <- rowSums(tempx)/100
temp2 <- rowSums(tempy)/108
temp3 <- rowSums(tempz)/93

Totrelab <- cbind(temp1, temp2, temp3)
Totrelab <- cbind(temp1, temp2)
Totrelab["Lactobacillus_crispatus/acidophilus", ] <- Totrelab["Lactobacillus_crispatus/acidophilus", ] + Totrelab["Lactobacillus_crispatus", ]
Totrelab[rownames(Totrelab) != "Lactobacillus_crispatus", ]
  o <- order( rowSums(Totrelab) , decreasing = T)
  #sample_o <- names(group_annotation)[ order( group_annotation )]
  temp_a <- Totrelab[o,]*100
  write.csv(temp_a, "totalRelAbundance.csv")

  # temp <- temp[-grep("NA",rownames(temp)),]
  temp_b <- rbind(temp_a[1:19,], Other = colSums(temp_a[20:nrow(temp_a),]) )

layout(matrix(c(2) , ncol= 2,byrow = T),heights = c(1))
#par(mar=c(2,7))
layout(matrix(c(2) , ncol= 2,byrow = T),
       heights = c(1))
  par(mar=c(2,7))
  barplot( temp_a, las=2 , border = NA ,ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp_a),levels = rownames(temp_a))] ,names.arg = rep("",ncol(temp_a)))
  legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],title.adj = 0,
         legend = rownames(temp_a),xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
          title="top abundant bacteria",
         pt.bg = taxa_pal)
```

```{r fig1, eval=FALSE, fig.cap="Figure 1. Box plot of the total lengths of live and dead sparrows following a snowstorm in Providence", fig.height=5, fig.width=12, message=FALSE, warning=FALSE, include=FALSE}

#############
# LODA DATA #
#############
#ALL_DATA <- readRDS("~/Box/Projects/J_Mjosberg/NBIS_housekeeping_normalization/analysis_all_tissues/all_ilc_dataset.rds")
#pal <- c(scales::hue_pal()(8),RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2") )

########################
# DEFINE FIGURE LAYOUT #
########################
layout(matrix(c(1,2,5,
                1,2,6,
                1,2,7,
                1,2,8,
                1,2,9,
                1,3,10,
                1,3,11,
                1,3,12,
                1,3,13,
                1,3,14,
                1,4,15,
                1,4,16,
                1,4,16,
                17,4,16,
                17,4,16),
              nrow = 15, byrow = T), 
   widths=c(3,1.5,1.5), heights=c(1,1))
#layout.show(16)

################
# PLOT FIGURES #
################
figlabels <- letters

### A
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "RNA_snn_res.3", label = T, frame=F, main="",cex = .7)
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = "", col = "white", pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Clustering")

### B
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Tissue", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = levels(ALL_DATA$Tissue), col = pal[1:3], pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Tissue")

### C
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Celltype", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = levels(ALL_DATA$Celltype), col = pal[1:4], pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Celltype")

### D
################
par(mar = c(0,0,0,0), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot_meta(ALL_DATA,red = "umap",feat = "Donor", label = T, frame=F, main="")
fig_label(figlabels[1],cex = 3,font = 2,region = "plot") ; figlabels <- figlabels[-1]
legend("topright", legend = "", col = "white", pch = 16, bty="n",
       pt.cex = 1,cex = 1,title = "Donor")

### E
################
my_gene_list <- c("PTGDR2","TNFSF10","TMEM273","KRT1","HPGD","FAM110A","ZP1","PKIB","TESPA1","CSGALNACT1")
par(mar=c(0.5,6,0,.5))
for(i in my_gene_list){ 
try({violins(data = ALL_DATA,gene = i,clustering = "Celltype",assay = "RNA",plot_y_axis = F,plot_x_axis = F,ylab="",main="",bw = 1)
mtext(side = 2, text = i, line = 1,las=1,cex = .8)})
}
mtext(side = 1, at=1:5, text = sort(unique(ALL_DATA@meta.data[,"Celltype"])), line = 1, las=1)
text(1:5, par("usr")[3], labels = sort(unique(ALL_DATA@meta.data[,"Celltype"])), srt = 45, adj = c(1,0.5), xpd = T, cex=1)

### F
################
my_image <- image_read("images/sparrow.jpg")
par(mar = c(1, 1, .5, .5), mgp = c(3, 0.5, 0),cex.lab = 1, cex.main = 1.2)
plot(my_image)
fig_label(figlabels[1],cex = 3,font = 2,region = "figure") ; figlabels <- figlabels[-1]

### G
################

```