### 2021-11-11 reports/rmarkdown/manuscript/Figure1.Rmd
Fig.1 minor changes

---
title: "Figure 1.Luminal Patient Grouping"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(openxlsx)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
group_annotation <- factor(setNames(metadata$Luminal_gr,metadata$ID))

SNN_patients <- read.csv("../../../results/patient_SNN_graph.csv",row.names = 1)
bac_communities <- read.csv2("../../../results/bacterial_communities.csv",row.names = 1)

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- colnames(TRX)[colSums(TRX)!=0]

#################
# COLOR PALETTS #
#################
lvl <- c("L. iners", "Gardnerella", "L. crispatus/acidophilus","Prevotella", "Atopobium","Sneathia", "L. jensenii", "Megasphaera", "Streptococcus", "Anaerococcus", "BVAB2", "Escherichia/Shigella", "BVAB1", "Dialister", "Mycoplasma", "Bifidobacterium", "Other")

pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
#show_col(pal)
```

```{r Figure 1., echo=FALSE, fig.asp=1, fig.height=7, fig.pos="t", fig.width=6, message=FALSE, warning=FALSE}
figlabels <- letters

layout(matrix(c(1,1,1,1,1,1,
                2,2,2,2,2,2,
                3,3,3,3,3,3,
                5,6,7,7,8,8,
                5,4,7,7,8,8,
                5,4,7,7,8,8) , ncol= 6,byrow = T),
                               widths = c(1.87,1.42,.7,.8,.8,.7,.8),
                               heights = c(1.6,.65,1.5,.95,.9,.2))
### A
#################################
# COHORT SCHEMATIC ILLUSTRATION #
#################################
par(mar=c(0,3.3,0,1))
fig1a <- magick::image_read("../../../supplementary_files/Scematic figure.pdf", density = 900 )
plot(c(0,1),c(0,1),axes=F,type="n",xlab="",ylab="",xaxs="i",yaxs="i")
rasterImage(fig1a, 0, 0, 1, 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### B
#############################
# DIVERSITY + TAX ABUNDANCE #
#############################

ASV <- list(c("Luminal_gr","ASV_CVL_V3_normalized_batch_corrected"),
          c("Tissue_gr","ASV_tissue_V3_normalized_batch_corrected")
         )
div_list <- list()
for(i in ASV){
  x <- datasets_all_samples[[i[2]]]
  gr <- factor(setNames(metadata[[i[1]]],metadata$ID))

  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","Simpson","Inv Simpson")

  div <- tibble(ID=names(shann),shann, simp, invsimp ) %>% rename_with(., ~paste0(c("ID","Shannon","Simpson","Inv Simpson")))
  div_list[[length(div_list) + 1]] <- div

  if(i[2] == "ASV_tissue_V3_normalized_batch_corrected"){break}

  par(mar=c(0,5,1.5,1.5)) #b,l,t,r
  barlist( data = divers[,order(group_annotation)],
         main = "", xlab="", #bg="white", #cex=.6, 
         genes = c("Shannon","Simpson","Inv Simpson"),
         draw_mean_lines=F,col = pal[gr[order(group_annotation)]])

  if(i[2] == "ASV_CVL_V3_normalized_batch_corrected"){
    end <- table(pal[factor(group_annotation[order(group_annotation)])])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(colnames(divers)))-.2 )
    start <- c(0.5, map_dbl(end[1:4], ~.x+0.7))

    axis(1, at = c(start, end), label = F, pos =3.3 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(3.3, 3.3), xpd=T, cex=.8))
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(gr), ~text(x=.x, y=3.8, .y,xpd=T))
  }

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  par(mar=c(1,5,1,1.5)) #b,l,t,r
  temp[is.nan(temp)] <- 0

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- names(group_annotation)[ order( group_annotation )]
  temp <- temp[o,sample_o]*100
  temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=gsub("_gr","",i[1]),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = lvl)] ,names.arg = rep("",ncol(temp)))

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2, bg="white",
        col= c("tomato","orange","#d7d7d7")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-2,labels = "      BV",cex=.5,xpd=T)

  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-8,bg="white",
        col= c("#d7d7d7","tomato")[as.numeric(metadata$HIVstatus)][order( group_annotation )],
        pch=15,cex=.7,xpd=T)
  text(par("usr")[2],par("usr")[3]-8,labels = "       HIV",cex=.5,xpd=T)

}

###########
# LEGENDS #
###########
par(mar=c(1,0,0,.5)) #b,l,t,r
empty_plot()
legend(x = -0.04, y = 1.04, 
       legend = c("L1 L. crispatus/jensenii", "L2 L. iners", "L3 Gardnerella", "L4 High diverse", "L5 Other") ,
       xjust = 0,yjust = 1,title.adj = 0,  ncol =1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,title = "Study Groups",
       pt.bg = pal)

par(mar=c(0,3.3,1.5,.5)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[1:12],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)

par(mar=c(0,0,1.5,1)) #b,l,t,r
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],
         legend = lvl[13:17],xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal[13:17])

### C + D
####################
# PATIENT GROUPING #
####################

### Louvain ###
g <- graph_from_adjacency_matrix(as.matrix(SNN_patients), mode = "undirected",diag = F,weighted = T)
set.seed(1)
l <- layout_with_fr(g,niter=3000,start.temp=30)

par(mar=c(0,0,1,1))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10, layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "Louvain", line = -2, cex.main = 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

### UMAP ###
all_microbiome <- datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]]
all_microbiome <- all_microbiome[ rowSums(all_microbiome>0) >= 3 ,  ]
cors <- cor(all_microbiome[,], method = "pearson")
adj <- (1-cors)/2
U <- uwot::umap(adj,n_neighbors = 30)
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] , 
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,layout=U,
      vertex.size=10,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
title(main = "UMAP", line = -2, cex.main = 1)

#add label
add_letter(figlabels[1]); figlabels <- figlabels[-1]

```

```{r Suppl. Table 1., include=FALSE}
# Total Relative Abundance 
raw <- c("ASV_CVL_V3", "ASV_tissue_V3") %>% set_names()
raw_list <- raw %>%
  map(., ~read_csv(paste0("../../../results/",.x,"_raw_counts.csv")))  %>%
  map(., ~column_to_rownames(., var = "X1")) %>%          
  map(., ~dplyr::select(., any_of(sample_use)))

r <- raw_list %>%
  map(., ~mutate(.,across(where(is.numeric), ~ ./sum(.))) ) %>%
  map(., ~(rowSums(.x)/length(colnames(.x)) )*100 ) %>%
  bind_rows(., .id="ID") %>%
    pivot_longer(-ID) %>% 
    pivot_wider(names_from = ID, values_from = value) %>%
    mutate_at(vars(name), ~replace(., ASV_CVL_V3<0.5514392, "Other")) %>%
    group_by(name) %>%
    summarize(across(where(is.numeric), ~sum(.))) %>%
    arrange(-ASV_CVL_V3) %>%
    dplyr::rename(Luminal="ASV_CVL_V3",Tissue="ASV_tissue_V3") %>%
    rename(Taxa="name") %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 1)))

# Alfa Diversity Stats
div_stat <- div_list %>%
  set_names(map(ASV,1)) %>%
  map(., ~summarise_all(.x, list(median=median, IQR=IQR), na.rm = TRUE)) %>%
  map(., ~pivot_longer(.x, everything(), names_sep = "_", names_to = c(".value", "Stats"), )) %>%
  bind_rows(.id = "Sample type") %>%
  select(-ID) %>%
  mutate("Sample type" = gsub("_gr", "", .$"Sample type")) %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 2)))

# Bacterial Communities
bac_communities <- bac_communities %>% 
  rownames_to_column(var = "Taxa") %>%
  mutate("bacterial communities"= factor(paste0("BC",sprintf("%02d",bac_communities$x)))) %>%
  select(-x)

pages <- list("Total Relative Abundance"=r ,"Alfa Diversity"=div_stat, "Bacterial Communities"=bac_communities)
write.xlsx(pages, file=paste0("./Suppl.Tbl/","Suppl.Tbl.01",".xlsx"))

```