### 2021-11-12 reports/rmarkdown/manuscript/Figure4-5.Rmd

---
title: "DGE's and Enrichment"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
  library(readxl)
  library(enrichR)
  library(igraph)
  library(rafalib)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)
sample_use <- metadata$ID

gr_name <- c("Luminal", "Tissue")
out <- paste0("./Suppl.Tbl/")

DEG_files <- c("Suppl.Tbl.04 DEGs_Luminal.xlsx",
               "Suppl.Tbl.08 DEGs_Tissue.xlsx")
enrichR_files <- c("Suppl.Tbl.07 TF_PPI_Luminal.xlsx",
                   "Suppl.Tbl.11 TF_PPI_Tissue.xlsx")

##################
# DEFINE CUTOFFS #
##################
dge_cutoff <- 0.01 # p-value
enrich_cutoff <- 0.1 # p-value
no_genes_cutoff <- 3
col_name <- paste0("genes_in(pValue<",dge_cutoff,")")

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

## Heatmap DGEs and enrichment
```{r Across functions, message=FALSE, warning=FALSE, include=FALSE}
##################
# DEG's FUNCTION #
##################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
edgR_dge.fun <- function(group, confound) {

  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~groups, data=metadata[sample_use,]) # without confounders
  }else{
    c <- "_conf"
    design <- model.matrix(~groups+HIVstatus+Contraception,
                         data=metadata[sample_use,])
  }

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- DGEList(counts=TRX_counts[,sample_use], remove.zeros = T)
  y <- calcNormFactors(y,method = "TMM")
  y <- estimateGLMCommonDisp(y,design)
  y <- estimateGLMTagwiseDisp(y,design)
  fit <- glmFit(y, design)

  lrt <- glmLRT(fit,coef=2:n_gr) # with intercept
  top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
  colnames(top) <- sub(group,"",colnames(top))
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Genes")

return(list(design=design, y=y, fit=fit, lrt=lrt, top=top, c=c))
}

#######################
# ENRICHMENT FUNCTION #
#######################
# group_name <- "Luminal_gr"
# top <- dge_L$top
# y <- dge_L$y
# c <- dge_L$c
gene_mod_enrich.fun <- function(top, y, c, db, group_name, cuttoff){
  p_val <- pull(top, "PValue")
  top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dge <- top$Genes[ top_dge ]

  top_TRX_counts <- edgeR::cpm(y,normalized.lib.sizes = T,log = T)[top_dge, ]
  top_TRX_counts <- t(apply(top_TRX_counts,1,function(x){scale(x,T,T)}))
  top_TRX_counts[top_TRX_counts > 5] <- 5
  colnames(top_TRX_counts) <- colnames(TRX_counts)

  h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
  cuttoff <- cuttoff[names(cuttoff) == db]
  gene_module <- cutree(h, h=cuttoff)

  source("../enrichment_function.R")
  if(db == "GO"){
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")}else{
  gmt_list <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")}
  res_list <- lapply( unique( gene_module[h$order] ),function(x){
    temp <- compute_enrichment(genes = names(gene_module)[gene_module==x],
                              gmt_list = gmt_list,
                              min_terms_pathway = 10,
                              max_terms_pathway = 300,
                              min_overlap = 3,
                              sort_by_pvalue = T)
    temp <- temp[!grepl("REGULATION",rownames(temp)),]
    return(temp)
  } )

  return(list(res_list=res_list, top_TRX_counts=top_TRX_counts, 
              gene_module=gene_module, cuttoff=cuttoff, h=h, c=c))
}

```

```{r Pairwise functions, message=FALSE, warning=FALSE, include=FALSE}
###########################
# PAIRWISE DEG's FUNCTION #
###########################
pairwise_dge.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups),na.last = NA))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  #fit2 <- contrasts.fit(fit, CONTRASTS)
  #lrt_ <- glmLRT(fit2, coef = 2:5)
  lrt <- map(s, ~ glmLRT(fit, contrast = CONTRASTS[,.x]))
  top <- map(lrt, ~topTags(.x,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]) %>% map(., ~rownames_to_column(.x, var = "Genes"))

  results <- map_df(lrt, ~decideTests(.x, method="separate", p.value = 0.05))
  summary <- bind_cols( "FDR<0.05"= c("Down","NotSig","Up"), summary(results))
  colnames(summary) <- c("FDR<0.05", g)
  top <- c(summary = list(summary), top)

return(list(lrt=lrt, top=top, CONTRASTS=CONTRASTS))
}
################################
# PAIRWISE ENRICHMENT FUNCTION #
################################
enrich_list <- function(x, gmt_list){
  temp <- compute_enrichment(genes = x,
                            gmt_list = gmt_list,
                            min_terms_pathway = 10,
                            max_terms_pathway = 300,
                            min_overlap = 3,
                            sort_by_pvalue = T)
  temp <- temp[!grepl("REGULATION",rownames(temp)),]
  temp <- rownames_to_column(temp, var = "Terms")
  return(temp)
} 
```

```{r Across tables, fig.height=4, fig.width=7, warning=F, include=FALSE}

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
TRX_counts <- TRX_counts[,sample_use]

#########
# DEG's #
#########
# Replace missing values with average in sexwork_months column
m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
metadata <- metadata %>%
  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))

dge_L <- edgR_dge.fun("Luminal_gr", confound=T)
dge_T <- edgR_dge.fun("Tissue_gr", confound=T)

###############
# ENRICHMENT #
##############
db <- c("GO", "KEGG") %>% set_names()
enrich_L <- db %>%
  map(., ~gene_mod_enrich.fun(dge_L$top,dge_L$y,dge_L$c,
                              db = .x,
                              group_name = "Luminal_gr", 
                              cuttoff = c(GO=2.5, KEGG=2.5)#3
                              )) 
enrich_T <- db %>%
  map(., ~gene_mod_enrich.fun(dge_T$top,dge_T$y,dge_T$c,
                              db = .x,
                              group_name = "Tissue_gr",
                              cuttoff = c(GO=1.6, KEGG=1.6)
                              ))

```

```{r Pairwise tables, include=FALSE}
##################
# PAIRWISE DEG's #
##################
pairwise_dge_L <- pairwise_dge.fun(group = "Luminal_gr", 
                                   fit=dge_L$fit, 
                                   design=dge_L$design )
pairwise_dge_T <- pairwise_dge.fun(group = "Tissue_gr",
                                   fit=dge_T$fit,
                                   design=dge_T$design)

###############################
# PAIRWISE ENRICHMENT GO/KEGG #
###############################
source("../enrichment_function.R")
gmt_list_KEGG <- fgsea::gmtPathways("../../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
gmt_list_GO <- fgsea::gmtPathways("../../../supplementary_files/c5.bp.v6.2.symbols.gmt.txt")

#info_logFC <- map(top_dfs[[1]], ~table(abs(.x$logFC) > 1))
#info_PValue <- map(top_dfs[[1]], ~table(.$PValue < 0.01))

l_top <- list(Luminal = pairwise_dge_L$top[-1], Tissue = pairwise_dge_T$top[-1])
# up and down regulated genes seperatly
t_split <- modify_depth(l_top, 2, ~ .x %>%
        dplyr::filter(., .$PValue < 0.01) %>%
        #dplyr::filter(., abs(.$logFC) < 1) %>%
        mutate(., reg = ifelse(.$logFC > 0, "up", "down")) %>%
        split(., .$reg)
  )

## GO enrichment 
pairwise_GO <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_GO)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation")) 
## KEGG enrichment
pairwise_KEGG <- map_depth(t_split, 3, ~enrich_list(.x$Genes, gmt_list_KEGG)) %>%
  map_depth(., 2, ~bind_rows(.x, .id = "regulation"))

```

```{r Suppl. Tbl. 4 & 8, include=FALSE}
# DEG's across and pairwise
pages <- list(list(Across=dge_L$top), pairwise=pairwise_dge_L$top) %>% flatten()
write.xlsx(pages, file=paste0(out,DEG_files[1]))

pages <- list(list(Across=dge_T$top), pairwise=pairwise_dge_T$top) %>% flatten()
write.xlsx(pages, file=paste0(out,DEG_files[2]))
```

```{r Suppl. Tbl. 5,6 & 9,10, eval=FALSE, include=FALSE}
# Enrichment Across
df_across <- enframe(list(Luminal=enrich_L, Tissue=enrich_T)) %>%  
  unnest_longer(value) %>%
  mutate(res = map(.$value, "res_list")) %>%
  unnest_longer(res) %>%
  mutate(res = map(res, ~rownames_to_column(.x, var = "Terms"))) %>%
  mutate(res = map(res, ~filter(.x, pvalue <= 0.01))) %>%
  group_by(name, value_id) %>%
  mutate(mod = row_number()) %>% ungroup() %>%
  mutate(res = set_names(.$res, paste0(.$value_id,"_gene_module_",.$mod))) %>%
  group_split(name) %>% set_names(., map(., ~unique(.x$name)))

write.xlsx(df_across[[1]]$res, file=paste0(out,"Suppl.Tbl.05 Enrichment_Across_Luminal",".xlsx"))
write.xlsx(df_across[[2]]$res, file=paste0(out,"Suppl.Tbl.09 Enrichment_Across_Tissue",".xlsx"))

# Enrichment pairwise
df_pairwise <- enframe(list(GO=pairwise_GO, KEGG=pairwise_KEGG)) %>%
  unnest_longer(value) %>%
  unnest_longer(value, names_repair ="unique") %>%
  mutate(value = map(value, ~filter(.x, FDR <= 0.05))) %>%
  mutate(value = set_names(.$value, paste0(.$name,"_",.$value_id...3))) %>%
  group_split(value_id...4) %>% set_names(., map(., ~unique(.x$value_id...4)))

write.xlsx(df_pairwise[[1]]$value, file=paste0(out,"Suppl.Tbl.06 Enrichment_Pairwise_Luminal",".xlsx"))
write.xlsx(df_pairwise[[2]]$value, file=paste0(out,"Suppl.Tbl.10 Enrichment_Pairwise_Tissue",".xlsx"))
```

```{r TF PPI Analysis function, include=FALSE}
TF_PPI_analysis.fun <- function(out, filename, gr_name, outfile) {
  # Get DEG's files and filter p-value < 0.01
  sheets <- excel_sheets(paste0(out, filename)) %>% set_names(.) %>% .[-c(1:2)]
  detables <- map(sheets, ~read_xlsx(paste0(out, filename), sheet = .x, col_names = TRUE)) %>%
    map(.,~mutate(.x, DE = ifelse(.x$PValue > dge_cutoff , 0, ifelse(.x$logFC < 0, -1, 1 )))) %>% 
    map(.,~filter(., .$PValue <= 0.01))

    # Running enrichments for up- or down-regulated genes
    save_l <- list()
    for(i in sheets ){
      message(paste0("Processing comparisson   ", i))
      temp <- detables[[i]]
      temp <- temp[ !duplicated(temp$Genes) & !is.na(temp$Genes) , ]
      temp <- mutate(temp, rownames = Genes)
      temp <- column_to_rownames(temp, var = "rownames")

      save_f <- list()
      for(j in c(1,-1)){
        gene_list <- temp$Genes[ temp$DE == j ]

        #Transcription_Factor_PPIs
        f <- enrichr(genes = gene_list, databases = "Transcription_Factor_PPIs")[[1]]
        if (is.data.frame(f) && nrow(f)>0){
          f <- f[order(f$P.value,decreasing = F), -c(5:6)]
          f <- cbind("Transcription Factor"=f[["Term"]], setNames(list(length(gene_list)), col_name), f[,-c(1)])
          f <- f[f$P.value < enrich_cutoff & as.numeric(sapply(strsplit(f[,"Overlap"],"/"),"[[", 1)) >= no_genes_cutoff,]
          f_n <- set_names(list(f), paste0(ifelse(j==1,"up","down"))) 
          save_f <- append(save_f, f_n)

        }else{if(length(save_f) == 0){save_f <- append(save_f, f_n)} }
      }
      l <- bind_rows(save_f, .id = "Regulation") %>%
        filter(Adjusted.P.value <= 0.05)
      l <- set_names(list(l), paste0(i))
      save_l <- append(save_l, l ) 
    } 
    write.xlsx(save_l, file=paste0(out, outfile), row.names=FALSE)
  }
```

```{r TF PPI Analysis, eval=FALSE, include=FALSE}
TF_PPI_analysis.fun(out, DEG_files[1], gr_name[1], enrichR_files[1]) # writes Suppl.Tbl.07
TF_PPI_analysis.fun(out, DEG_files[2], gr_name[2], enrichR_files[2]) # writes Suppl.Tbl.11
```

```{r plot functions, warning=FALSE, include=FALSE}
############################
# GO/KEGG HEATMAP FUNCTION #
############################
# sample_use <- metadata$ID
# gr = "Tissue_gr"
# name = "Tissue"
# enrich = enrich_T
# db <- "GO"
# db <- "KEGG"
# plot_heatmap.fun(enrich_T, db, "Tissue_gr", "Tissue")
plot_heatmap.fun <- function(enrich, db, gr, name, N=3, Layout){
  figlabels <- letters
  df <- na.omit(metadata[sample_use, c(gr, "ID")])
  df <- arrange(df, df[1])
  samples <- df[, "ID"]
  names(samples) <- df[, gr]

  conf <- pluck(enrich, db, "c")
  h <- pluck(enrich, db, "h")
  cuttoff <- pluck(enrich, db, "cuttoff")
  res_list <- pluck(enrich, db, "res_list")
  top_TRX_counts <- pluck(enrich, db, "top_TRX_counts")[,samples]
  gene_module <- pluck(enrich, db, "gene_module")

  if(missing(Layout)){
    layout(matrix(c(1,2,3,4,5,
                  1,2,3,4, 5),
                  nrow = 2,ncol = 5,byrow = T),widths = c(1,3.5,.36,8,8))
  }else{layout(Layout[[1]],Layout[[2]],Layout[[3]])}

  ## Heatmap
  par(mar=c(4,.1,2.5,.1))
  plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
  title(xlab="Height", line=2)
  abline(v=cuttoff,xpd=F,col="red",lty=2)
  points( rep(-.1,length(gene_module)),
          seq(length(gene_module),1,length.out = length(gene_module)),
          col=taxa_pal[factor(gene_module[h$order])],
          pch=15,cex=.5,xpd=F)

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("           Differentialy Expressed Genes", side=3, las = 1, cex = .6, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

  end <- table(pal[factor(names(samples))[ order(df[1])]])
  end <- end[order(factor(names(end), levels = pal))]
  end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
  start <- c(0, map_dbl(end[1:4], ~.x+0.015))

  axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
  map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
  # text labels
  t <- map2(start, end, ~(.x+.y)/2)
  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

  ## Meta info bars
  points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
          col=pal[factor(names(samples))[ order(df[1])]],
          pch=15,xpd=T,cex=.7)
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
          col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45) 
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
          col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45)

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )

  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))
  # pvalues <- unlist(lapply(res_list,function(x){ x$jaccard_index [1:N] }))
  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  terms <- gsub("endoplasmic reticulum", "ER", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""

  #pvalues <- setNames(pvalues,paste0(terms," (",genes,")"))
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
  plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
               use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",
               yaxs="i",plot_weights = F)

  text(par("usr")[1],par("usr")[3]-.008,labels = "    BV",cex=.6, xpd=T)
  text(par("usr")[1],par("usr")[3]-.028,labels = "    HIV",cex=.6, xpd=T)

  # KEGG
  if(name == "Tissue")
  {par(mar=c(4,.1,2.5,13))}else #bottom, left, top, right  
      {par(mar=c(4,.1,2.5,8))}
  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(p-value)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)

  # GO
  res_list <- pluck(enrich, "GO", "res_list")
  #top_TRX_counts <- pluck(enrich, "GO", "top_TRX_counts")[,samples]

  ## Gene module enrichment
  names(res_list) <- unique( gene_module[h$order] )
  pvalues <- unlist(lapply(res_list,function(x){ -log10(x$pvalue) [1:N] }))

  pvalues[is.na(pvalues)] <- 0
  terms <- unlist(lapply(res_list,function(x){ rownames(x) [1:N] }))
  terms[is.na(terms)] <- ""
  terms <- gsub("_", " ", terms)
  terms <- gsub("(GO |KEGG )|(\\w+)", "\\1\\L\\2", terms, perl = TRUE)
  terms <- gsub("endoplasmic reticulum", "ER", terms, perl = TRUE)
  terms <- gsub("establishment of ", "", terms, perl = TRUE)
  genes <- unlist(lapply(res_list,function(x){ x$genes [1:N] }))
  genes[is.na(genes)] <- ""
  pvalues <- setNames(pvalues,paste0(terms,""))
  module_color <- unlist(lapply(names(res_list),function(x){ rep(x,N) }))

  barplot( rev(pvalues) , horiz = T , col= taxa_pal[factor(rev(module_color))],
           xaxs="i",yaxs="i",xlim=c(0,1.2*max(pvalues)),ylab="",las=1,
           names.arg = "",xpd = FALSE)
  abline(v=2,xpd=F,col="red",lty=2)
  points( rev(pvalues), seq(1,length(pvalues) )*1.2-.5,pch=21,
          bg=taxa_pal[factor(rev(module_color))] )
  title(xlab="-log10(p-value)", line=2)
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        adj=0,cex=.9,xpd=T,rev(gsub("\\(\\)","",names(pvalues)) ))
  text( rev(pvalues)+strwidth("L") , seq(1,length(pvalues) )*1.2-.5, 
        labels = rev(sub(" [(].*","",names(pvalues)) ), adj=0,cex=.9,
        col = paste0(taxa_pal[factor(rev(module_color))],90) ,xpd=T)
}

###########################
# ASSEMBLE PLOTS FUNCTION #
###########################
arrange_plots.fun <- function(degs, group, gr_name, comparison, enrichment_db, file, Layout){

  ### A
  ###################
  # GO/KEGG HEATMAP #
  ###################
  plot_heatmap.fun(degs, "KEGG", group, gr_name, Layout=Layout)

  ### B
  #####################
  # TF PPI ENRICHMENT #
  #####################
  # Create plot for each enrichment
  n <- 10
  for( x in comparison ){
  for( i in enrichment_db ){
    DD <- Inf
    UU <- 0
    names_DD <- ""
    names_UU <- ""

    f_DOWN <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "down")
    f_UP <- read_xlsx(paste0(out, file), sheet = x) %>% filter(Regulation == "up")

    if (is.data.frame(f_DOWN) && nrow(f_DOWN)>0){
      tempD_e <- f_DOWN
      DD <- log10(tempD_e$P.value)[1:min(n,length(tempD_e$P.value))]
      names_DD <- (as.character(tempD_e$"Transcription Factor"))[1:min(n,length(tempD_e$P.value))]
    }
    if (is.data.frame(f_DOWN) && nrow(f_DOWN)>0){
      tempU_e <- f_UP
      UU <- rev(-log10(tempU_e$P.value)[1:min(n,length(tempU_e$P.value))])
      names_UU <- rev(as.character(tempU_e$"Transcription Factor")[1:min(n,length(tempU_e$P.value))])
    }

    comb <- c( DD/abs(max(abs(DD))) , UU/abs(max(abs(UU))) )
    comb <- replace_na(comb, 0)

    if (identical(comb, c(0,0))){next}

    par(mar = c(1.2, 4.7, 2.2, 1))  #bottom, left, top, right
    barplot( c(rep(NA, each=20-length(comb)), comb),  
             xlim=c(-1.1,1.1),
             # xlim=c(min(c(comb,-2/abs(max(abs(DD))) ))*1.1, 
             #                          max(c(comb,2/abs(max(abs(DD)))) )*1.1), 
                        cex.names = .9, #mar=c(1,25,2,2), #main=paste(x, i, sep="\n"),
             names.arg = c(rep("", each=20-length(comb)),names_DD, names_UU ), 
             horiz = T,las= 1, border=NA,xaxt='n',line=-.2, #xlab="-log10 p-value",
             col = c(rep(NA, each=20-length(comb)), 
                     rev(colorRampPalette( c("firebrick3","gray","navy") )(19))[(comb+1)*9+1])
    )
    title(main=paste(x, str_replace_all(i, "_"," "), sep="\n"), line = 0.1, cex.main=.9)

    if(x == "L1-L2" | x == "T1-T2"){add_letter("b")}
    points_y <- rev(seq(par("usr")[4]-1.45, by=-1.2, length=length(c(UU,DD))))
             points( comb , points_y,
                     #mar=c(1,20,2,2),
                     bg=colorRampPalette( c("navy","gray","firebrick3") )(19)[(comb+1)*9+1],
                     pch=21, col="black",cex=1.5)
             lines(x=c(0, 0), y= c(points_y[1]-.6, par("usr")[4]-1)) 
             lines(x=c(-2/abs(max(abs(DD))),-2/abs(max(abs(DD)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3)) # pval=0.01
             lines(x=c(2/abs(max(abs(UU))),2/abs(max(abs(UU)))), 
                   y= c(points_y[1]-.6, par("usr")[4]-1), lty=c(3))  # pval=0.01
             text(0,points_y[1]-1.5, "-log10 p-value", xpd=T)
    }
  }
}

dev.copy2pdf(file=paste0("./test.pdf"),
    width = 6.6,
    height = 6, paper = "a4"
    )
```

```{r Figure 4., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
# Layout
l <- list(mat = matrix(c(1,2,3,4,4,4,5,5,5,
                         1,2,3,4,4,4,5,5,5,
                         6,6,6,7,7,8,8,9,9,
                         10,10,10,11,11,12,12,13,13),
                nrow = 3,ncol = 9,byrow = T),
          widths = c(1.14,3.5,.36,2.5,2.5,2.5,2.5,2.5,2.5),
          heigth = c(1,1,1.5)
          )
#layout.show(layout(l[[1]],l[[2]],l[[3]]))

sheets <- excel_sheets(paste0(out, DEG_files[1])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_L, "Luminal_gr", "Luminal", sheets, "TF_PPI_enrichment", enrichR_files[1], Layout = l)
```

```{r Figure 5., message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
# comparison <- sheets
# enrichment_db <- "TF_PPI_enrichment"
# degs <- enrich_T
# group <- "Tissue_gr" 
# gr_name <- "Tissue"

# Layout
l <- list(mat = matrix(c(1,2,3,4,4,4,5,5,5,
                         1,2,3,4,4,4,5,5,5,
                         6,6,6,7,7,8,8,9,9,
                         10,10,10,11,11,12,12,13,13),
                nrow = 3,ncol = 9,byrow = T),
          widths = c(1.14,3.5,.36,2.5,2.5,2.5,2.5,2.5,2.5),
          heigth = c(1,1,1.5)
          )

sheets <- excel_sheets(paste0(out, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)
```

```{r UMAP, eval=FALSE, fig.height=2.5, fig.width=3, include=FALSE}
# , fig.height=3, fig.width=3.5
###############
#  TRX UMAP   #
###############
plot_UMAP.fun <- function(top_TRX_counts, group){
  #par(mar=c(4,4,2,6)) #b,l,t,r
  par(mar=c(2.5,2.5,2,3)) #b,l,t,r
  #par(mar=c(10,12,2,12)) #b,l,t,r
  set.seed(1)
  UMAP_TRX <- uwot::umap(t(top_TRX_counts),n_neighbors = 10,
                   metric = "correlation",min_dist = 0.1,spread = 5,
                   negative_sample_rate = 10)
  plot(UMAP_TRX,bg=pal[factor(group)],main="UMAP on top DGE RNAseq",
       pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
  mtext("UMAP1", side=1, line=0.7, cex=0.8)
  mtext("UMAP2", side=2, line=0.7, cex=0.8)
  legend(par("usr")[2],par("usr")[4],title.adj = 0,
         #legend = levels(factor(metadata$Luminal_gr)),
         legend = sort(unique(group)),
         xjust = 0,yjust = 1, cex=0.7,
         bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
}

plot_UMAP.fun(enrich_L$GO$top_TRX_counts, metadata$Luminal_gr)
m <- filter(metadata, !is.na(Tissue_gr)) 
plot_UMAP.fun(enrich_T$GO$top_TRX_counts[,m$ID], m$Tissue_gr)

#plot_UMAP.fun(enrich_T$GO$top_TRX_counts, metadata$Tissue_gr)
dev.copy2pdf(file=paste0("./UMAP",".pdf"),
    width = 4,
    height = 3.5, paper = "a4" #units = "cm", res = 300 
    )

```

 <!-- -->