### 2022-02-24 reports/02_data_preprocessing.Rmd
move normalization to seperate file

---
title: "Merge sequencing runs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(results    = "hold",
                      message    = FALSE,
                      warning    = FALSE)

#setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports")
```

# Load libraries and other scripts

```{r Load data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
library(tidyverse)
})  })  })

#########
# PATHS #
#########
#input_dir <- "../../Gabriella_repo/data/"
input_dir <- "../results/01_taxonomy_improved_output/"
result_dir <- "../results/02_data_preprocessing_output/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
dataset_names <- c("ASV_tissue_V3_B2.csv", # Tissue, Boston run 2 (95 samples)
                   "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
                   #"ASV_CVL_V2_B1.csv",   # CVL V2, Boston run 1 (27 samples)
                   #"ASV_CVL_V2_B2.csv",   # CVL V2, Boston run 2 (49 samples)
                   #"ASV_CVL_V2_S.csv",    # CVL V2, CTMR         (62 samples)
                    "ASV_CVL_V3_B1.csv")   # CVL V3, Boston run 1 (111 samples)

datasets <- map(dataset_names, ~read.csv(paste0(input_dir,.x))) %>% set_names(., dataset_names)
metadata <- read.csv(paste0("../data/", "metadata.csv"),row.names = 1)
trx <- read.csv(paste0("../data/","Raw_gene_counts_matrix.csv"),row.names = 1)

map(datasets, ~dim(.x))
```

```{r Functions}
# create "other" taxonomy
other_taxa.fun <- function(df) {
  other_taxa <- df %>%
  unite("taxa_other", Kingdom:Species, sep = ";", remove = F, na.rm = T) %>%
  mutate(taxa_other = ifelse(is.na(.$Species), paste0(.$taxa_other, ";other"), .$taxa_other)) %>%
  #mutate(taxa_other = str_replace(.$taxa_other, ";NA.+|;NA", ";other")) %>%
  mutate(taxa_other = str_extract(.$taxa_other, "([^;]+);([^;]+)$")) %>% # get the two last taxa levels
  mutate(taxa_other = sub('^(.*);(other)', '\\2;\\1', .$taxa_other)) # place "other" first
return(other_taxa)
}

# create genus lacto taxonomy
genus_lacto.fun <- function(df){
  genus_lacto <- df %>%
    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste("L.", "crispatus/acidophilus"), 
                                ifelse(grepl("reuteri",.$Species), paste("L.", "reuteri/oris/frumenti/antri"),
                                  ifelse(grepl("gasseri",.$Species), paste("L.","gasseri/johnsonii/taiwanensis"),
                                    ifelse(grepl("murinus",.$Species), paste("L.","murinus/animalis/apodemi/salivarius"),
                                      ifelse(grepl("plantarum",.$Species), 
                                           paste("L.", "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
                                        ifelse(grepl("Lactobacillus",.$Genus), paste("L.", .$Species), 
                                              as.character(.$Genus_lacto))))))) )
  return(genus_lacto)
}

aggregate.fun <- function(df, level){
  level <- enquo(level)
  aggregated <- df %>%
    dplyr::rename(Taxonomy = !!level) %>%
    unite(., "Full_taxonomy", Kingdom:Species, sep = ";") %>%
    dplyr::select(-any_of(c("Sequence", "Seq_ID", "Genus_lacto", "taxa_other", "Full_taxonomy"))) %>%
    group_by(Taxonomy) %>%
    summarise(across(where(is.numeric), sum)) %>%
    ungroup()
return(aggregated)
}
```

# Taxonomy agglomeration

```{r Agglomerate Taxa}
temp <- map(datasets, ~.x  %>%
            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
            other_taxa.fun(.) %>% # create "other" taxonomy 
            genus_lacto.fun(.)  # create genus lacto taxonomy
            ) %>%  
        {. ->> temp_ } %>% 
        map(., ~aggregate.fun(., "Genus_lacto")) # agglomerate on genus/lacto spp. lvl.

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_, .id = "dataset") %>% select(1:12) %>% unique() 
#write.csv(seq_taxa, paste0(result_dir,"SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )

#arrange all datasets to have the same rownames with all bacteria
common_microbes_fun <- function(x) {
  t <- bind_rows(temp) %>%
    select(Taxonomy) %>%
    unique() %>%
    left_join(x, by="Taxonomy") %>%
    mutate(across(Taxonomy, ~replace_na(.x, "Not assigned"))) %>%
    mutate(across(-Taxonomy, ~replace_na(.x, 0))) %>%
    column_to_rownames(var = "Taxonomy")
  return(t)
}
datasets <- map(temp, ~common_microbes_fun(.x))
lapply(datasets,dim)

```

# Merging and renaming

```{r merge sequencing runs}
names(datasets)

#No batch correction needed in tissue_V3 dataset, only add it to the other samples
datasets[["ASV_tissue_V3"]] <- cbind( datasets[["ASV_tissue_V3_B2.csv"]],datasets[["ASV_tissue_V3_B1.csv"]] )

datasets <- datasets[-c(1,2)]
lapply(datasets,dim)

```

```{r save raw microbiome files, fig.height=5, fig.width=5}

datasets <- lapply(datasets,function(x){
  x <- x[,sort(colnames(x))]
  return(x)
})

# raw files
n <- c("ASV_Luminal", "ASV_Tissue")
map2(datasets, n,  ~write.csv(.x, file=paste0(result_dir,.y,"_raw_counts.csv"),row.names = T) )

```