### 2022-02-24 reports/manuscript/Figure6.Rmd

---
title: "Figure 6. Differential Protein expression"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6.6929133858,
  fig.path="./Figures/",
  dev=c('pdf', 'png'), dpi=300,
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(readxl)
  library(rstatix)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#########
# PATHS #
#########
result_dir <- "./Suppl.Tbl/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI_Bradley_et.al.xlsx")
cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)

#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

```{r DPE Functions, message=FALSE, warning=FALSE, include=FALSE}
################################
# DEG's FUNCTION ACROSS GROUPS #
################################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
#group <- "L_T_groups"
limma_dpe.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~  groups, data=metadata[sample_use,]) } # without confounders
  else{
    c <- "_conf"
    # design <- model.matrix(~groups+HIVstatus+Contraception+age+sexwork_months,
    #                      data=metadata[sample_use,])
    design <- model.matrix(~ groups+Contraception,
                         data=metadata[sample_use,])
  }
  #design <- model.matrix(~0 + groups+Contraception+HIVstatus+age+sexwork_months, 
                         #data=metadata[sample_use,]) # without intercept

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- Protein[,sample_use]
  fit <- lmFit(y, design)
  fit2 <- contrasts.fit(fit, coefficients=2:n_gr)
  efit <- eBayes(fit2)

  # DPE across all groups
  top <- topTable(efit, adjust.method = "BH",number = "all") # ANOVA
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Antibody")

return(list(design=design, y=y, fit=fit, efit=efit, top=top, c=c))
}

###########################
# DEG's FUNCTION PAIRWISE #
###########################
pairwise_dpe.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups)))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  # DPE pairwise comparison
  fit2 <- contrasts.fit(fit, CONTRASTS)
  efit <- eBayes(fit2)
  top <- map(s, ~ topTable(efit, coef=.x, adjust.method = "BH",number = "all",sort.by = "P")) %>%
    map(., ~rownames_to_column(.x, var = "Antibody"))

  results <- decideTests(fit2, method="separate", p.value = 0.05)
  summary <- rownames_to_column(as.data.frame.array(summary(results)), var="FDR<0.05")
  top <- c(summary = list(summary), top)

  #saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
  #saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
return(list(efit=efit, top=top, CONTRASTS=CONTRASTS))
}

```

```{r format data, echo=FALSE, warning=FALSE, message=FALSE}
to_remove <- c("CSTB_HPA017380","SPINK5_HPA009067","SPRR3_HPA024330","FGA_HPA064755","S100A12_HPA002881","KRT1_HPA019797","KRT1_HPA062908","LGMN_HPA001426","TACSTD2_HPA043104","FLNA_HPA001115","GPX3_HPA071520","CSTA_HPA001031","SERPINB5_HPA019132","ITIH2_HPA059150","NCF2_HPA002327","ANXA3_HPA013431")

Protein <- Protein_V3 %>%
  select(ID, contains("HPA")) %>%
  filter(ID %in% metadata$ID) %>%
  pivot_longer(., -ID, names_to = "Antibody") %>%
  pivot_wider(names_from = "ID", values_from = "value" ) %>%
  filter(!(Antibody %in% to_remove)) %>%
  column_to_rownames(var = "Antibody")

```

```{r Across and Pairwise tables, include=FALSE}
# Samples to use
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
sample_use <- intersect(metadata$ID, colnames(TRX)[colSums(TRX)!=0])
sample_use <- intersect(colnames(Protein), sample_use)

# Across groups DPE
dpe_L <- limma_dpe.fun("Luminal_gr", confound=T)
dpe_T <- limma_dpe.fun("Tissue_gr", confound=T)

# L2 to T2/T3
metadata <- metadata %>% 
  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))

dpe_LT <- limma_dpe.fun("L_T_groups", confound=T)

# Pairwise DPE's
pairwise_dpe_L <- pairwise_dpe.fun("Luminal_gr", 
                                    fit=dpe_L$fit, 
                                    design=dpe_L$design )
pairwise_dge_T <- pairwise_dpe.fun("Tissue_gr",
                                   fit=dpe_T$fit,
                                   design=dpe_T$design)

```

```{r heatmap functions, echo=FALSE}
top_dpe.fun <- function(top){
  p_val <- pull(top, "P.Value")
  top_dpe <- (p_val < 0.05) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dpe <- top$Antibody[ top_dpe ]

  top_PROT_mfi <- Protein[top_dpe, sample_use]

  return(top_PROT_mfi)
}

# top_PROT_mfi <- t
# gr <- "Luminal_gr"
# plot_heatmap.fun(t, "Luminal_gr", scale = T)
plot_heatmap.fun <- function(top_PROT_mfi, gr, scale = FALSE, layout = "none"){
  df <- na.omit(metadata[sample_use, c(gr, "ID")])
  df <- arrange(df, df[1])
  samples <- df[, "ID"]
  names(samples) <- df[, gr]

  if(scale == TRUE) {
    n <- colnames(top_PROT_mfi)
    top_PROT_mfi <- t(apply(top_PROT_mfi,1,function(x){scale(x,T,T)}))
    colnames(top_PROT_mfi) <- n
    breaks = seq(-5,5,length.out = 100)
  }else{breaks = seq(0,5,length.out = 100)}

  h <- hclust( as.dist( (1- cor(t(top_PROT_mfi)))/2 ), method = "ward.D2")

  if(layout == "multi") {
    layout(matrix(c(1,2,5,6,
                    7,7,8,8,
                    3,3,4,4),
           nrow = 3, ncol = 4,byrow = T),
           widths = c(1,2,.8,2), heights = c(1,1,.01) ) }
  if(layout == "single") {
    layout(matrix(c(1,2),
                    nrow = 1, ncol = 2,byrow = T),widths = c(1,2) ) }

    ## Dendogram
    par(mgp=c(0,0.2,0), mar=c(4,.6,2.5,.2)) #(bottom, left, top, right) 
    plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .7 )
    #title(xlab="height", line=3)
    mtext( c("Height"), side=1, las = 1, cex = .7, line =1,xpd=T)

    if(gr == "Tissue_gr") {
      height <-1.5 
      add_letter("c")
    }else{height <- round(max(h$height))
      add_letter("a")}
    #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob

    ## Heatmap
    par(mar=c(4,.2,2.5,2.5)) #(bottom, left, top, right) 
    image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
           col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
           breaks = breaks,
           axes=F)
    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, font = 2, line =1.5,xpd=T)
    mtext( c("Study Groups"), side=1, las = 1, cex = .7, line = 1,xpd=T)

    # Study group annotation
    end <- table(pal[factor(names(samples))[ order(df[1])]])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
    start <- c(0, map_dbl(end[1:4], ~.x+0.015))
    end <- c(end[1:length(end)-1], par("usr")[2]-.005)

    axis(1, at = c(start, end), label = F, pos =1.032 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(1.032, 1.032), xpd=T, cex=.8))
    # text labels
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

    # Boxes 
    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,-.002,-.002),
            y=c(.488,.488,1.008, 1.008,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

    ## Meta info bars
    if(gr == "Luminal_gr"){size = .5}else{size=.6}
    points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
            col=pal[factor(names(samples))[ order(df[1])]],
            pch=15,xpd=T,cex=.7)

    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.008,
            col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
            pch=15,xpd=T,cex=size) 
    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
            col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
            pch=15,xpd=T,cex=size)

    mtext( paste0(" ", gsub("_.+","",rownames(top_PROT_mfi)[rev(h$order)])), side=4, las = 1, cex = .45,
           at=seq(0,1,length.out = length(rownames(top_PROT_mfi))), xpd=T)

    mtext( c(" BV"," HIV"), side=4, las = 1, cex = .45,
           at=c(par("usr")[3]-.008,par("usr")[3]-.022), xpd=T)
}

```

```{r sig. cytokines function, include=FALSE}
sig_stars.fun <- function(df_krusk, df_dunn, group, n = 6, order = NULL){
  samples_use <- metadata[,c(group, "ID")] %>% na.omit() %>% filter(ID %in% colnames(cytokine_data))

  # select only sig. cytokines 
  sig <- df_krusk %>% filter(p.adj<0.05) %>% pull(Cytokines)
  sig_df <- cytokine_data_log[sig,samples_use$ID]

  groupings <- samples_use[[group]] [ match( colnames(sig_df[,1:ncol(sig_df)-1]) , as.character(samples_use$ID)  ) ]
  groupings <- c(groupings)

   if(is.null(order)){
    g <- rownames(sig_df)
  }else{g <- order}

  stars_df <- df_dunn %>%
    mutate(x = (as.integer(str_extract(.$group1, "\\d")) + 
                  as.integer(str_extract(.$group2, "\\d")))/2) %>%
    mutate(line_x0 = as.integer(str_extract(.$group1, "\\d"))) %>%
    mutate(line_x1 = as.integer(str_extract(.$group2, "\\d"))) %>%
    mutate(y = (as.numeric(factor(Cytokines, levels = rev(g))))-.3) %>%
    arrange(group1, group2) %>%
    dplyr::group_by(Cytokines) %>%
    nest() %>%
    mutate(y_ = map(data, ~accumulate(.$y, ~.x + .1))) %>%
    unnest(c(data, y_))

  return(list(sig_df=sig_df, stars_df=stars_df, groupings=groupings, samples_use=samples_use))
}
```

```{r Cytokine stats tests, include=FALSE}
cytokine_data$ID <- metadata [ match( cytokine_data$PatID , as.character(metadata$PatID)  ) ,"ID" ]
cytokine_data <- cytokine_data[ !is.na(cytokine_data$ID) , ]
rownames(cytokine_data) <- cytokine_data$ID
cytokine_data <- cytokine_data[,grepl("CVL",colnames(cytokine_data))]
colnames(cytokine_data) <- sub("CVL_","",colnames(cytokine_data))
cytokine_data <- t(cytokine_data)
cytokine_data[is.na(cytokine_data)] <- 0.01
cytokine_data <- log2(cytokine_data+1)
cytokine_data_log <- cytokine_data

cyto2 <- as.data.frame(t(cytokine_data))
cyto2$ID<-row.names(cyto2)

#select data and re-format to long format
cytogr.long <- metadata %>%
  select(ID, Luminal_gr, Tissue_gr) %>%
  left_join(.,cyto2, by="ID") %>%
  pivot_longer(c(-Luminal_gr, -Tissue_gr, -ID), names_to = "Cytokines", values_to = "value")

test.krusk_dunn.fun <- function(df, var, group) {
  var <- enquo(var)

  stat.test.krusk <- df %>%
  group_by(!!(var)) %>%
  kruskal_test(., as.formula(paste("value", "~", group))) %>%
  adjust_pvalue(method = "BH") %>%
  select(-.y., -n, -df) %>%
  arrange(p) %>%
  rename(p.value = "p") %>%
  add_significance()

# select the significant proteins from kruskal
sig <- stat.test.krusk %>% filter(p.adj<0.05) %>% pull(!!var)

### Post hoc test
stat.test.dunn <- df %>%
  filter(!!var %in% sig) %>%
  group_by(!!var) %>%
  dunn_test(., as.formula(paste("value", "~", group))) %>%
  #dunn_test(value ~ !!group) %>%
  select(-.y.) %>%
  arrange(p) %>%
  rename(p.value = "p") %>%
  filter(p.adj<0.05)

  sig2 <- stat.test.dunn[stat.test.dunn$p.adj<0.05,]

 l <- list(stat.test.krusk, stat.test.dunn)
 n <-  c("Kruskal", "Dunn's") %>% paste0(., "_", group, sep = "")
 l <- set_names(l, n)

return(l)
}

n <- c("Luminal_gr", "Tissue_gr") %>% set_names(c("Luminal", "Tissue"))
stat <- imap(n, ~test.krusk_dunn.fun(cytogr.long, Cytokines, .x))
s <- flatten(stat)
```

```{r Suppl. Tbl. 15 16 & 17, include=FALSE}
names <- c("15 Luminal", "16 Tissue") %>% set_names()

# save TopTables pairwise
l_top_p <- list(pairwise_dpe_L$top, pairwise_dge_T$top)  %>% set_names(names)

# save TopTables across + cytokines
l_top_a <- list( list(Across = dpe_L$top), list(Across = dpe_T$top))  %>% set_names(names)
l <- list(l_top_a, stat, l_top_p)
p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))

imap(p_list, ~write.xlsx(.x,file=paste0(result_dir,"Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0(result_dir,"Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
```

```{r Figure 6, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
figlabels <- letters
####################################
# TOP DIFF. EXPRESSED PROT HEATMAP #
####################################
### A
# Luminal groups
t <- top_dpe.fun(dpe_L$top)
plot_heatmap.fun(t, "Luminal_gr", scale = T, layout = "multi")

### C
# Tissue groups
t <- top_dpe.fun(dpe_T$top)
par(mfg = c(2,1)) # adds the remaning plots to the previous layout
plot_heatmap.fun(t, "Tissue_gr", scale = T)

##############################
# SIG. CYTOKINES VIOLIN PLOT #
##############################
### B
# Luminal groups
g <- c("IL1a", "MIG", "IP10", "IL1b", "IL12_p70", "MCP1")
Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g)
stars <- Luminal$stars_df

par( mar=c(1.4,5,1.5,2)) #b,l,t,r 
violist(data=Luminal$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= g, #add_ylims = T, #gives starting value of y-axis and max value of all data points
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings))
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("b")

### D
# Tissue groups
Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
stars <- Tissue$stars_df

par( mar=c(20.4,6,1.5,3)) #b,l,t,r
violist(data=Tissue$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= rownames(Tissue$sig_df), 
          col = pal, transparency = 50, #add_ylims = T,
          clustering=Tissue$groupings )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("d")

```

**Figure 6. Protein Heatmap**. **a** and **c** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively. (p-value < 0.05) **b** and **d** sig. cytokines for Luminal and Tissue respectively.

```{r All dpe plot, eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
library(cowplot)

plot_heatmap.fun(Protein[,sample_use], "Luminal_gr")
p.plot <- recordPlot()

plot_heatmap.fun(Protein[,sample_use], "Tissue_gr")
p.plot_T <- recordPlot()

# side by side unscaled
plot_grid(ggdraw(p.plot),ggdraw(p.plot_T), 
          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F)

```