### 2022-08-25 reports/manuscript/Suppl.Figure3-4.Rmd

---
title: "Suppl. Figure 3. & 4. Mirobiome-KEGG correlations."
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, out_dir, ...) {
    source("../../code/knit_function.R");
    custom_knit(inputFile, "../../lab_book/SupplFigure3_4/", ...)
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path="./Suppl.Figures/",
  fig.align = "center",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(openxlsx)
  library(fgsea)
  library(rafalib)
  library(Rcpp)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#########
# PATHS #
#########
PATH <- "../../.."
reprod_files <- "../../data/reprod.Suppl.Fig.3_4/"

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
SNN_bacteria <- read.csv("../../results/04_clustering_output/bacteria_SNN_graph.csv",row.names = 1)
# setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
#datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
#SNN_bacteria <- read.csv("../../../results/bacteria_SNN_graph.csv",row.names = 1)
bac_communities <- read.xlsx("./Suppl.Tbl/Suppl.Tbl.01.xlsx", sheet = "Bacterial Communities")

```

```{r eval=FALSE, include=FALSE}
counts <- t(as.data.frame( lapply(datasets_all_samples,colSums) ))
counts[is.na(counts)] <- 0
```

```{r Suppl.Fig.3 Bacterial drivers, warning=FALSE, fig.width=12, fig.height=13}
myletter <- letters
#reprod. <- FALSE
reprod. <- TRUE

# ASV <- list(c("Luminal","ASV_CVL_V3_normalized_batch_corrected"),
#             c("Tissue","ASV_tissue_V3_normalized_batch_corrected")
#             )
ASV <- list(c("Luminal","ASV_Luminal_normalized"),
            c("Tissue","ASV_Tissue_normalized")
            )
# i <- ASV[[1]]
for(i in ASV){
  cat(paste0("Correlation between RNAseq and ",i[1], " 16S","\n"))

  # because hclust is non reproducible 
  if(reprod. == TRUE){
    NESes <- read.csv2(paste0(reprod_files,i[1],"_TRX_Bact_norm_enrichment_scores",".csv"), row.names = 1)
  }else{

    TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
    bac_dataset <- datasets_all_samples[[i[2]]]
    samples_TRX_and_CVL3 <- colnames(counts) [ colSums(counts[c("Tissue_RNAseq_V3_normalized",i[2]),] > 0)==2 ]

    #filter TRX dataset
    TRX <- TRX [ , samples_TRX_and_CVL3 ]
    TRX <- TRX [ rowSums(TRX>0)>= 2 , ]
    # only top 5000 most variabe genes are included in the correlation
    top_vars_TRX <- names(sort(apply(TRX,1,var),decreasing = T)[1:5000] )
    TRX <- TRX[ top_vars_TRX , ]
    dim(TRX)

    #filter bac_dataset dataset
    bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
    # keeping taxa that is present in at least two samples :
    bac_dataset <- bac_dataset [ rowSums(bac_dataset>0)>= 2 , ]
    top_vars_bacs <- names(sort(apply(bac_dataset,1,var),decreasing = T)[1:100] )
    dim(bac_dataset)

    cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )

    bacteria_use <- colnames(cors)
    gmt <- gmtPathways(paste0(PATH, "/supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt"))

    enrichments <- lapply(bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
      res <- fgsea(pathways = gmt,stats = sort(cors[,x],decreasing = T),nper=10000)
      return(res)
    })
    names(enrichments) <- bacteria_use

    pvalues <- lapply(enrichments,function(x) setNames(x$pval,x$pathway) )
    pvalues <- t(as.data.frame(pvalues))
    pvalues <- -log10( pvalues )
    top_pathways <- names(sort(apply(pvalues,1,median),T))[1:50]

    NESes <- lapply(enrichments,function(x) setNames(x$NES,x$pathway) )
    NESes <- t(as.data.frame(NESes))
    NESes[is.na(NESes)] <- 0
    # replace all NESes values with 0 if they don't have a significant p-value < 0.05:
    NESes[ pvalues < -log10(0.05) ] <- 0 
    # filter out bacterium and pathways with <10 significant NES scores 
    NESes <- NESes[rowSums(NESes!=0) >= 10 , colSums(NESes!=0) >= 10]

    set.seed(1)
    o_kegg <- hclust( as.dist( (1-cor(NESes))/2 ),"ward.D2")$order
    o_bacs <- hclust( as.dist( (1-cor(t(NESes) ))/2 ),"ward.D2")$order
    NESes <- NESes[o_bacs, o_kegg]

    bact <- gsub("\\.{1}", "/", rownames(NESes))
    bact <- gsub("\\/{2}", ". ", bact)
    rownames(NESes) <- bact

    #write.csv2(NESes, paste0("../../../results/",i[1],"_TRX_Bact_norm_enrichment_scores",".csv"))

  }

  terms <- gsub("_", " ", colnames(NESes))
  terms <- gsub("(GO |KEGG )", "", terms, perl = TRUE)
  colnames(NESes) <- terms

  ### Suppl. Figure 3-4
  #######################################
  # FUNCTIONAL ASSOCIATION BACT AND TRX #
  #######################################
  par(mfrow=c(1,1),mar=c(12,0,2,4)) #b,l,t,r
  image( t(NESes[nrow(NESes):1,]),col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-3,3,length.out = 92),
         axes=F,border=NA,main="tissue RNAseq KEGG pathways",xlab="",ylab=i,line=.4,cex.main=1,font.main=1)
  text(  par("usr")[c(4)] , seq(1,0,length.out = nrow(NESes)),
        labels = rownames(NESes), srt = 0, adj = c(0,.5), xpd = TRUE, cex=.4)
  text( seq(0,1,length.out = ncol(NESes)) , par("usr")[c(1)],
        labels = colnames(NESes), srt = 90, adj = c(1,.5), xpd = TRUE, cex=.4)

}

```

**Suppl. Figure 3-4.** Heatmaps showing the functional association of microbiome datasets with the expression profiles in the RNAseq dataset. Briefly, bacterial abundances from each dataset were correlated with the gene expression of the top 5000 highly variable genes from the RNAseq dataset, generating a correlation matrix between bacteria and genes. Then for each bacteria, we rank genes based on their correlation to that bacteria and perform gene set enrichment anlaysis (GSEA) using the KEGG gene annotation database. This, in turn, will result in a matrix associating every bacteria with every KEGG process in the tissue. The heatmap shows the normalized enrichment score (NES). Only enrichments with pvalue below 0.05 are shown.Only bacterium and pathways with at least 10 significant NES scores were included in the heatmap.

```{r eval=FALSE, include=FALSE}
g <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
cl <- igraph::cluster_louvain(g)
cors <- cor(t(datasets_all_samples[[i]][rowSums(datasets_all_samples[[i]] > 0)>=10,]))

module_means <- bac_communities %>%
  left_join(as_tibble(datasets_all_samples[[i]], rownames = "Taxa"), by="Taxa")  %>%
  group_by(bacterial.communities) %>%
  summarise(across(where(is.numeric), ~sum(.x))) %>%
  mutate(across(where(is.numeric), ~.x/c(table(bac_communities$bacterial.communities))) ) %>%
  select(-bacterial.communities) %>%
  #column_to_rownames(var="bacterial.communities") %>%
  as.matrix() %>%
  `rownames<-`(unique( bac_communities$bacterial.communities)) 
bact_df <- module_means

all_microbiome <- cbind(datasets_all_samples[["ASV_tissue_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V3_normalized_batch_corrected"]],
                        datasets_all_samples[["ASV_CVL_V2_normalized_NOT_batch_corrected"]])
#cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )

module_means_per_dataset <- lapply(datasets_all_samples[[i]],function(x){
  return( rowsum( x[bac_communities$Taxa,] , group = bac_communities$bacterial.communities ) / c(table(bac_communities$bacterial.communities)) )
})

module_means <- bac_communities %>%
  left_join(as_tibble(datasets_all_samples[[i]], rownames = "Taxa"), by="Taxa")  %>%
  group_by(bacterial.communities) %>%
  summarise(across(where(is.numeric), ~sum(.x))) %>%
  mutate(across(where(is.numeric), ~.x/c(table(bac_communities$bacterial.communities))) )

b  <-  datasets_all_samples[[i]][bac_communities$Taxa,]
rowsum(datasets_all_samples[[i]][bac_communities$Taxa,] , group = bac_communities$bacterial.communities )
```

```{r eval=FALSE, fig.height=8, fig.width=10, include=FALSE}
myletter <- letters

NESse_list <- list()
pvalues_list <- list()

for(i in c("ASV_tissue_V3_normalized_batch_corrected",
           "ASV_CVL_V3_normalized_batch_corrected")){

  cat(paste0("\nCorrelation between RNAseq and ",sub("_norm.*","",i)))

  TRX <- as.matrix(datasets_all_samples[["Tissue_RNAseq_V3_normalized"]])
  bac_dataset <- module_means_per_dataset[[i]]
  samples_TRX_and_CVL3 <- colnames(bac_dataset)

  # mypar(4,3,mar=c(4,2,2,2))
  #filter TRX dataset
  TRX <- TRX [ , samples_TRX_and_CVL3 ]
  TRX <- TRX [ rowSums(TRX>0)>= 5 , ]

  #filter bac_dataset dataset
  bac_dataset <- bac_dataset [ , samples_TRX_and_CVL3 ]
  dim(bac_dataset)

  cors <- cor( t(rbind(TRX) ) ,  t(rbind(bac_dataset) ) )
  dim(cors)

  bacteria_use <- colnames(cors)
  #gmt <- gmtPathways("../../supplementary_files/h.all.v6.2.symbols.gmt.txt")
  gmt <- gmtPathways("../../supplementary_files/c2.cp.kegg.v6.2.symbols.gmt.txt")
  cl <- parallel::makePSOCKcluster(parallel::detectCores()-1)
  invisible(parallel::clusterEvalQ(cl, {c("gmt","cors");library(fgsea);library(stats);library(base)}))
  enrichments <- parallel::parLapply(cl,bacteria_use,gmt=gmt,cors=cors,function(x,gmt,cors){
    res <- fgseaMultilevel(pathways = gmt,stats = cors[,x])
    return(res)
  })
  parallel::stopCluster(cl)
  names(enrichments) <- paste0( "bac.com.", bacteria_use )

  allpathways <- unique(unlist(lapply(enrichments,function(x){x[,1]})))
  allpvals <- setNames(rep(NA,length(allpathways)),allpathways)

  pvalues <- lapply(enrichments,function(x){
    temp <- allpvals
    temp[x$pathway] <- x$pval
    return(temp)
  })
  pvalues <- t(as.data.frame(pvalues))
  pvalues <- -log10( pvalues )
  pvalues[is.na(pvalues)] <- 0

  allNES <- setNames(rep(NA,length(allpathways)),allpathways)
  NESes <- lapply(enrichments,function(x) {
    temp <- allNES
    temp[x$pathway] <- x$NES
    return(temp)
  })
  NESes <- t(as.data.frame(NESes))
  NESes[is.na(NESes)] <- 0
  # NESes[ pvalues < -log10(0.1) ] <- 0

  NESse_list[[i]] <- NESes
  pvalues_list[[i]] <- pvalues

}

write.csv2(NESse_list,"../../results/Hallmark_NESse_list.csv")
write.csv2(pvalues_list,"../../results/Hallmark_pvalues_list.csv")

common_pathways <- unique( unlist( lapply( NESse_list , function(x) colnames(x)[colSums(x!=0)>=1] ) ) )

layout(matrix(c(1,2,3,4,4),nrow = 1))
par(mar=c(6,1,2,0))
for( i in names(NESse_list)){
  x <- NESse_list[[i]][,common_pathways]
  image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
         axes=F,border=NA,main=sub("_norm.*","",i),xlab="",ylab="",line=.4,cex.main=1,font.main=1)
  # add_letter(myletter); myletter <- myletter[-1]
  text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
      labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
}
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)])
text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      labels = sub("HALLMARK_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

cat("\\clearpage")

# Combined representation of the two datasets
layout(matrix(c(1,2,2,2,2),nrow = 1))
par(mar=c(6,1,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rev(rownames(x)), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
empty_plot(xaxs="i",yaxs="i",
           xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main="HALLMARK")
text(  par("usr")[c(1)],seq(0,1,length.out = ncol(x)) ,
      labels = rev(sub("HALLMARK_","",colnames(x))), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

# Combined representation of the two datasets
layout(matrix(c(1,1,1,1,1,2,2),nrow = 1))
par(mar=c(6,30,3,0))
x <- (NESse_list[[1]][,common_pathways] + NESse_list[[2]][,common_pathways]) / 2
x[NESse_list[[1]][,common_pathways] == 0] <- 0
x[NESse_list[[2]][,common_pathways] == 0] <- 0
x <- x[, colSums(x!=0)>0]

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
non_zeros <- colSums(TRX)!=0
TRX <- TRX [  , non_zeros ]
pathway_means <- t(sapply( colnames(x) , function(i){
  i <- gmt[[i]]
  i <- as.matrix( TRX[ rownames(TRX) %in% i , ] )
  if(ncol(i)>1){
    i <- ( i - apply(i,1,min) ) / ( apply(i,1,max)- apply(i,1,min) )
    i <- colMeans(i)
  } else {
    i <- (i - min(i)) / (max(i) - min(i))
  }
  return( i - min(i) )
} ))
plot_dots(data=pathway_means,main="patient groups",cex.main=1,font.main=1,
        genes=rownames(pathway_means),
        clustering=factor(metadata$joint_clustering[non_zeros]),
        srt=0,pal=c("blue","blue","navy","navy","grey95","grey95","firebrick3","black"))

par(mar=c(6,2,3,2))
image( x[,ncol(x):1],col=colorRampPalette(c("navy","grey95","firebrick"))(91),breaks=seq(-5,5,length.out = 92),
       axes=F,border=NA,main="CVL_v3 + tissue_v3\ncombined",xlab="",ylab="",line=.4,cex.main=1,font.main=1)
# add_letter(myletter); myletter <- myletter[-1]
text(   seq(0,1,length.out = nrow(x)) , par("usr")[c(3)] ,
    labels = rownames(x), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)
# empty_plot(xaxs="i",yaxs="i",
#            xlim=par("usr")[c(1,2)],ylim=par("usr")[c(3,4)],main=" ")
# text(  par("usr")[c(1)],seq(1,0,length.out = ncol(x)) ,
      # labels = sub("KEGG_","",colnames(x)), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)

```