### 2023-03-08 reports/manuscript/Figure6.Rmd
reorder cytokine plot for proof

---
title: "Figure 6. Differential Protein expression"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, out_dir, ...) {
    source("../../code/knit_function.R");
    custom_knit(inputFile, "../../lab_book/Figure6/", ...)
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results    = "hold",
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 6.6929133858,
  fig.align  = "center",
  fig.path   = "./Figures/",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r Load-data, message=FALSE, warning=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(readxl)
  library(rstatix)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#########
# PATHS #
#########
result_dir <- "./Suppl.Tbl/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

prot_url <- "https://doi.org/10.1371/journal.ppat.1010494.s017"
#cyt_url <- ""
#meta_url <- ""

############
# DOWNLOAD #
############
if(isFALSE(file.exists("../../data/Protein_norm_MFI.xlsx"))){download.file(prot_url, paste0("../../data/", "Protein_norm_MFI.xlsx"), method="auto")}
#if(isFALSE(file.exists("../../data/Cytokine_data.xlsx"))){download.file(cyt_url, paste0("../../data/", "Cytokine_data.xlsx"), method="auto")}
#if(isFALSE(file.exists("../../data/metadata.xlsx"))){download.file(meta_url, paste0("../../data/", "metadata.xlsx"), method="auto")}

#############
# LODA DATA #
#############
# datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI.xlsx", skip = 2) %>% rename(ID="Patient/Sample ID")
# cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)

cytokine_data <- read_xlsx(paste0("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/data/", "Cytokine_data.xlsx"), na = "NA", sheet = "Cytokine_data") 
#################
# COLOR PALETTS #
#################
pal <- c( "#0072B2", "#009E73","#D55E00", "#CC79A7", "#E69F00", "#999999")
taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
```

```{r DPE-Functions, message=FALSE, warning=FALSE, include=FALSE}
################################
# DEG's FUNCTION ACROSS GROUPS #
################################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
#group <- "L_T_groups"
limma_dpe.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

  if(any(is.na(groups))) {
    groups <- replace_na(groups, "X")
  }

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~  groups, data=metadata[sample_use,]) } # without confounders
  else{
    c <- "_conf"
    # design <- model.matrix(~groups+HIVstatus+Contraception+age+sexwork_months,
    #                      data=metadata[sample_use,])
    design <- model.matrix(~ groups+Contraception,
                         data=metadata[sample_use,])
  }
  #design <- model.matrix(~0 + groups+Contraception+HIVstatus+age+sexwork_months, 
                         #data=metadata[sample_use,]) # without intercept

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("\\(Intercept\\)","Intercept",colnames(design))
  colnames(design) <- sub(" ","_",colnames(design))

  y <- Protein[,sample_use]
  fit <- lmFit(y, design)
  fit2 <- contrasts.fit(fit, coefficients=2:n_gr)
  efit <- eBayes(fit2)

  # DPE across all groups
  top <- topTable(efit, adjust.method = "BH",number = "all") # ANOVA
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Antibody")

return(list(design=design, y=y, fit=fit, efit=efit, top=top, c=c))
}

###########################
# DEG's FUNCTION PAIRWISE #
###########################
pairwise_dpe.fun <- function(group, fit, design) {
  groups <- metadata[sample_use, group]
  levels <- as.character(sort(unique(groups)))

  # without intercept:
  n <- combn(levels,2) 
  g <- map_chr(seq_along(1:ncol(n)), ~paste(n[1,.x], n[2,.x], sep = "-"))
  s <- seq_along(1:ncol(n)) %>% set_names(g)
  #name <- map_chr(seq_along(1:ncol(n)), ~paste(str_extract(n[1,.x], "^g."), str_extract(n[2,.x], "^g."), sep = "_"))

  # with intercept:
  g <- sub(paste0("^(",levels[1], ")-(.*)"), '-\\2', g)

  CONTRASTS <- makeContrasts( contrasts=g,
                              levels = design )

  # DPE pairwise comparison
  fit2 <- contrasts.fit(fit, CONTRASTS)
  efit <- eBayes(fit2)
  top <- map(s, ~ topTable(efit, coef=.x, adjust.method = "BH",number = "all",sort.by = "P")) %>%
    map(., ~rownames_to_column(.x, var = "Antibody"))

  results <- decideTests(fit2, method="separate", p.value = 0.05)
  summary <- rownames_to_column(as.data.frame.array(summary(results)), var="FDR<0.05")
  top <- c(summary = list(summary), top)

  #saveRDS(top, file = "./TopTable_all_groups_06.04.2021.RDS")
  #saveRDS(top, file = "./TopTable_ModifiedM4_pairwise_17.05.2021.RDS")
return(list(efit=efit, top=top, CONTRASTS=CONTRASTS))
}

```

```{r format-data, echo=FALSE, warning=FALSE, message=FALSE}
to_remove <- c("CSTB_HPA017380","SPINK5_HPA009067","SPRR3_HPA024330","FGA_HPA064755","S100A12_HPA002881","KRT1_HPA019797","KRT1_HPA062908","LGMN_HPA001426","TACSTD2_HPA043104","FLNA_HPA001115","GPX3_HPA071520","CSTA_HPA001031","SERPINB5_HPA019132","ITIH2_HPA059150","NCF2_HPA002327","ANXA3_HPA013431")

Protein <- Protein_V3 %>%
  select(ID, contains("HPA")) %>%
  filter(ID %in% metadata$ID) %>%
  pivot_longer(., -ID, names_to = "Antibody") %>%
  pivot_wider(names_from = "ID", values_from = "value" ) %>%
  filter(!(Antibody %in% to_remove)) %>%
  column_to_rownames(var = "Antibody")

```

```{r Across-and-Pairwise-tables, include=FALSE}
# Samples to use
# TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
# sample_use <- intersect(metadata$ID, colnames(TRX)[colSums(TRX)!=0])
sample_use <- metadata$ID
sample_use <- intersect(colnames(Protein), sample_use)

# Across groups DPE
dpe_L <- limma_dpe.fun("Luminal_gr", confound=T)
dpe_T <- limma_dpe.fun("Tissue_gr", confound=T)

# L2 to T2/T3
metadata <- metadata %>% 
  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))

dpe_LT <- limma_dpe.fun("L_T_groups", confound=T)

# Pairwise DPE's
pairwise_dpe_L <- pairwise_dpe.fun("Luminal_gr", 
                                    fit=dpe_L$fit, 
                                    design=dpe_L$design )
pairwise_dge_T <- pairwise_dpe.fun("Tissue_gr",
                                   fit=dpe_T$fit,
                                   design=dpe_T$design)

```

```{r heatmap-functions, echo=FALSE}
top_dpe.fun <- function(top){
  p_val <- pull(top, "P.Value")
  top_dpe <- (p_val < 0.05) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dpe <- top$Antibody[ top_dpe ]

  top_PROT_mfi <- Protein[top_dpe, sample_use]

  return(top_PROT_mfi)
}

# top_PROT_mfi <- t
# sacale = T
# gr <- "Luminal_gr"
# plot_heatmap.fun(t, "Luminal_gr", scale = T)
plot_heatmap.fun <- function(top_PROT_mfi, gr, scale = FALSE, layout = "none"){
  df <- na.omit(metadata[sample_use, c(gr, "ID")])
  df <- arrange(df, df[1])
  samples <- df[, "ID"]
  names(samples) <- df[, gr]

  if(scale == TRUE) {
    n <- colnames(top_PROT_mfi)
    top_PROT_mfi <- t(apply(top_PROT_mfi,1,function(x){scale(x,T,T)}))
    colnames(top_PROT_mfi) <- n
    breaks = seq(-5,5,length.out = 100)
  }else{breaks = seq(0,5,length.out = 100)}

  h <- hclust( as.dist( (1- cor(t(top_PROT_mfi)))/2 ), method = "ward.D2")

  if(layout == "multi") {
    layout(matrix(c(1,2,5,6,
                    7,7,8,8,
                    7,7,9,9,
                    3,3,4,4),
           nrow = 4, ncol = 4,byrow = T),
           widths = c(1,2,.8,2), heights = c(1,.35,.35,.01) )
    # layout.show(9)
    }
  if(layout == "single") {
    layout(matrix(c(1,2),
                    nrow = 1, ncol = 2,byrow = T),widths = c(1,2) ) }

    ## Dendogram
    par(mgp=c(0,0.2,0), mar=c(4,.6,2.5,.2)) #(bottom, left, top, right) 
    plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .8 )
    #title(xlab="height", line=3)
    mtext( c("Height"), side=1, las = 1, cex = .7, line =1,xpd=T)

    if(gr == "Tissue_gr") {
      height <-1.5 
      add_letter("c")
    }else{
      # add_scale_legend(x = par("usr")[1]-0.1, height = 1.5,#diff(par("usr")[3:4])-20,
      #                pal = colorRampPalette(c("navy","grey95","firebrick"))(91),
      #                labels = c("-5","0","5"))
      # height <- round(max(h$height))
      # add_letter("a")
      }
    #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob

    ## Heatmap
    par(mar=c(4,.2,2.5,3.2)) #(bottom, left, top, right) 
    image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
           col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
           breaks = breaks,
           axes=F)
    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, font = 2, line =1.5,xpd=T)
    mtext( c("Study Groups"), side=1, las = 1, cex = .7, line = 1,xpd=T)

    # Study group annotation
    end <- table(pal[factor(names(samples))[ order(df[1])]])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
    start <- c(0, map_dbl(end[1:4], ~.x+0.015))
    end <- c(end[1:length(end)-1], par("usr")[2]-.005)

    axis(1, at = c(start, end), label = F, pos =1.032 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(1.032, 1.032), xpd=T, cex=.8))
    # text labels
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

    # Boxes 
    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,-.002,-.002),
            y=c(.488,.488,1.008, 1.008,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

    ## Meta info bars
    if(gr == "Luminal_gr"){size = .5}else{size=.6}
    points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
            col=pal[factor(names(samples))[ order(df[1])]],
            pch=15,xpd=T,cex=.7)

    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.008,
            col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
            pch=15,xpd=T,cex=size) 
    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.020,
            col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
            pch=15,xpd=T,cex=size)

    mtext( paste0(" ", gsub("_.+","",rownames(top_PROT_mfi)[rev(h$order)])), side=4, las = 1, cex = .53,
           at=seq(0,1,length.out = length(rownames(top_PROT_mfi))), xpd=T)

    mtext( c(" BV"," HIV"), side=4, las = 1, cex = .45,
           at=c(par("usr")[3]-.008,par("usr")[3]-.022), xpd=T)
}

```

```{r sig-cytokines-function, include=FALSE}
sig_stars.fun <- function(df_krusk, df_dunn, group, n = 6, order = NULL){
  samples_use <- metadata[,c(group, "ID")] %>% na.omit() %>% filter(ID %in% colnames(cytokine_data))

  # select only sig. cytokines 
  sig <- df_krusk %>% 
    filter(p.adj<0.05) %>%
    pull(Cytokines)
  sig_df <- cytokine_data_log[sig,samples_use$ID]

  groupings <- samples_use[[group]] [ match( colnames(sig_df[,1:ncol(sig_df)-1]) , as.character(samples_use$ID)  ) ]
  groupings <- c(groupings)

   if(is.null(order)){
    g <- rownames(sig_df)
  }else{g <- order}

  stars_df <- df_dunn %>%
    mutate(x = (as.integer(str_extract(.$group1, "\\d")) + 
                  as.integer(str_extract(.$group2, "\\d")))/2) %>%
    mutate(line_x0 = as.integer(str_extract(.$group1, "\\d"))) %>%
    mutate(line_x1 = as.integer(str_extract(.$group2, "\\d"))) %>%
    mutate(y = (as.numeric(factor(Cytokines, levels = rev(g))))-.3) %>%
    arrange(group1, group2) %>%
    dplyr::group_by(Cytokines) %>%
    nest() %>%
    mutate(y_ = map(data, ~accumulate(.$y, ~.x + .1))) %>%
    unnest(c(data, y_))

  return(list(sig_df=sig_df, stars_df=stars_df, groupings=groupings, samples_use=samples_use))
}
```

```{r Cytokine-stats-tests, include=TRUE}
cytokine_data <- cytokine_data %>% 
  column_to_rownames(var = "ID") %>% 
  as.matrix()

cytokine_data <- t(cytokine_data)
cytokine_data[is.na(cytokine_data)] <- 0.01
cytokine_data <- log2(cytokine_data+1)
cytokine_data_log <- cytokine_data

cyto2 <- as.data.frame(t(cytokine_data))
cyto2$ID<-row.names(cyto2)

#select data and re-format to long format
cytogr.long <- metadata %>%
  select(ID, Luminal_gr, Tissue_gr) %>%
  left_join(.,cyto2, by="ID") %>%
  pivot_longer(c(-Luminal_gr, -Tissue_gr, -ID), names_to = "Cytokines", values_to = "value")

test.krusk_dunn.fun <- function(df, var, group) {
  var <- enquo(var)

  stat.test.krusk <- df %>%
  group_by(!!(var)) %>%
  kruskal_test(., as.formula(paste("value", "~", group))) %>%
  adjust_pvalue(method = "BH") %>%
  select(-.y., -n, -df) %>%
  arrange(p) %>%
  rename(p.value = "p") %>%
  add_significance()

# select the significant proteins from kruskal
sig <- stat.test.krusk %>% filter(p.adj<0.05) %>% pull(!!var)

### Post hoc test
stat.test.dunn <- df %>%
  filter(!!var %in% sig) %>%
  group_by(!!var) %>%
  dunn_test(., as.formula(paste("value", "~", group))) %>%
  #dunn_test(value ~ !!group) %>%
  select(-.y.) %>%
  arrange(p) %>%
  rename(p.value = "p") %>%
  filter(p.adj<0.05)

  sig2 <- stat.test.dunn[stat.test.dunn$p.adj<0.05,]

 l <- list(stat.test.krusk, stat.test.dunn)
 n <-  c("Kruskal", "Dunn's") %>% paste0(., "_", group, sep = "")
 l <- set_names(l, n)

return(l)
}

n <- c("Luminal_gr", "Tissue_gr") %>% set_names(c("Luminal", "Tissue"))
stat <- imap(n, ~test.krusk_dunn.fun(cytogr.long, Cytokines, .x))
s <- flatten(stat)
```

```{r Suppl. Tbl. 15 16 & 17, include=TRUE, results = 'hide'}
names <- c("15 Luminal", "16 Tissue") %>% set_names()

# save TopTables pairwise
l_top_p <- list(pairwise_dpe_L$top, pairwise_dge_T$top)  %>% set_names(names)

# save TopTables across + cytokines
l_top_a <- list( list(Across = dpe_L$top), list(Across = dpe_T$top))  %>% set_names(names)
l <- list(l_top_a, stat, l_top_p)
p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))

imap(p_list, ~write.xlsx(.x,file=paste0(result_dir,"Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0(result_dir,"Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
```

```{r Figure 6, echo = TRUE, eval = TRUE, fig.height=10, fig.width=7}
# dev.new(height=10, width=6.6929133858, noRStudioGD = TRUE)
figlabels <- letters
####################################
# TOP DIFF. EXPRESSED PROT HEATMAP #
####################################
### A
# Luminal groups
t <- top_dpe.fun(dpe_L$top)
plot_heatmap.fun(t, "Luminal_gr", scale = T, layout = "multi")

### C
# Tissue groups
t <- top_dpe.fun(dpe_T$top)
par(mfg = c(2,1)) # adds the remaning plots to the previous layout
plot_heatmap.fun(t, "Tissue_gr", scale = T)

##############################
# SIG. CYTOKINES VIOLIN PLOT #
##############################
### B
# Luminal groups
g <- c("MIG", "IL1b", "IL12_p70", "MCP1", "IL1a", "IP10") 
#g <- c("MIG", "IL12_p70","IL1a", "IP10", "IL1b", "MCP1" )
lab <- c("MIG", "IL-1\u03b2", "IL-12-p70", "MCP-1", "IL-1\u03b1", "IP-10" )
#lab <- c("IP-10", "IL-1\u03b2", "IL-12-p70", "MCP-1","IL-1\u03b1", "MIG")
Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g[1:4])
rownames(Luminal$sig_df) <- lab
stars <- Luminal$stars_df

par( mar=c(1.4,5,1.5,2)) #b,l,t,r 
violist(data=Luminal$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= lab[1:4],
          #cex.axis = 1.1,
          #add_ylims = T, #gives starting value of y-axis and max value of all data points
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings))
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = 1.1)

add_letter("b")

Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g[5:6])
rownames(Luminal$sig_df) <- lab
stars <- Luminal$stars_df 
par(mar=c(1.4,5,1.5,2) ) #b,l,t,r
violist(data=Luminal$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= lab[5:6],
          #add_ylims = T, #gives starting value of y-axis and max value of all data points
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings))
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_)
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = 1.1)

### D
# Tissue groups
lab <- c("IL-1\u03b1", "IP-10")
Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
rownames(Tissue$sig_df) <- lab
stars <- Tissue$stars_df

#par(mar=c(20.4,6,1.5,3)) #b,l,t,r # ps = 20, point size
par(mar=c(1,5,1.5,5)) #b,l,t,r
violist(data=Tissue$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= rownames(Tissue$sig_df),
          col = pal, transparency = 50,
          # add_ylims = T,
          clustering=Tissue$groupings )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_)
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = 1.1)

add_letter("d")

```

**Figure 6. Characterization of the host protein profile as stratified by the luminal and tissue-adherent microbiome study groups.** All study samples were assessed for significant differences in protein levels across the study groups. **a**,**b** The five luminal study groups and **c**,**d** the five corresponding tissue groups. a, c Proteins with a p-value > 0.05 were omitted from the heatmap. Proteins with significantly different levels across the groups were clustered by hierarchical agglomerative clustering using inverse Pearson’s correlation as distance measure and Ward’s method (“ward. D2”) for linkage. Color-coded rows below the heatmaps show clinical diagnosis of bacterial vaginosis (BV, binned Nugent’s scores) and HIV diagnosis, respectively. Color coding for BV: Gray: negative, orange: intermediate, red: positive; and for HIV: Gray: HIV seronegative, red: HIV seropositive. b, d Violin plots of log2 transformed cytokine levels across the luminal and tissue study groups (as indicated on the x-axis), respectively. The asterisk and lines indicate statistically significant results of Dunn’s test with Benjamini Hochberg’s correction analysis. Adjusted p-values: * < 0.05, ** < 0.01, *** < 0.001, **** < 0.0001

```{r old-Figure 6, echo = FALSE, eval = FALSE, fig.height=10, fig.width=7}
# old layout
    # layout(matrix(c(1,2,5,6,
    #                 7,7,8,8,
    #                 3,3,4,4),
    #        nrow = 3, ncol = 4,byrow = T),
    #        widths = c(1,2,.8,2), heights = c(1,1,.01) )
figlabels <- letters
####################################
# TOP DIFF. EXPRESSED PROT HEATMAP #
####################################
### A
# Luminal groups
t <- top_dpe.fun(dpe_L$top)
plot_heatmap.fun(t, "Luminal_gr", scale = T, layout = "multi")

### C
# Tissue groups
t <- top_dpe.fun(dpe_T$top)
par(mfg = c(2,1)) # adds the remaning plots to the previous layout
plot_heatmap.fun(t, "Tissue_gr", scale = T)

##############################
# SIG. CYTOKINES VIOLIN PLOT #
##############################
### B
# Luminal groups
g <- c("IL1a", "MIG", "IP10", "IL1b", "IL12_p70", "MCP1")
lab <- c("IL-1\u03b1", "MIG", "IP-10", "IL-1\u03b2", "IL-12-p70", "MCP-1")
Luminal <- sig_stars.fun(s$Kruskal_Luminal_gr, s$`Dunn's_Luminal_gr`, "Luminal_gr", order = g)
rownames(Luminal$sig_df) <- lab
stars <- Luminal$stars_df

par( mar=c(1.4,5,1.5,2)) #b,l,t,r 
violist(data=Luminal$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= lab,
          #add_ylims = T, #gives starting value of y-axis and max value of all data points
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings))
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("b")

### D
# Tissue groups
lab <- c("IL-1\u03b1", "IP-10")
Tissue <- sig_stars.fun(s$Kruskal_Tissue_gr, s$`Dunn's_Tissue_gr`, "Tissue_gr")
rownames(Tissue$sig_df) <- lab
stars <- Tissue$stars_df

par(mar=c(20.4,6,1.5,3)) #b,l,t,r # ps = 20, point size
violist(data=Tissue$sig_df,
          srt=45, y_padding = 2, smooth = .2,
          genes= rownames(Tissue$sig_df), 
          col = pal, transparency = 50, 
          # add_ylims = T,
          clustering=Tissue$groupings )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("d")

```

```{r All-dpe-plot, eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
library(cowplot)

plot_heatmap.fun(Protein[,sample_use], "Luminal_gr")
p.plot <- recordPlot()

plot_heatmap.fun(Protein[,sample_use], "Tissue_gr")
p.plot_T <- recordPlot()

# side by side unscaled
plot_grid(ggdraw(p.plot),ggdraw(p.plot_T), 
          labels = c('a', 'b'),rel_widths=c(1,1), greedy = F)

```